---
title: "Bloodstream infections and associated mortality from CROs in Modena"
output:
  word_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
  pdf_document:
    toc: yes
---

# Introduction
This R notebook provides implementation codes for replicating data processing and analysis for examination of bloodstream infections (BSI) due to carbapenem resistant organisms (CROs) and equivalent carbapenem sensitivity pathogens, and competing risks of hospital stay and death


# 1. Preliminaries

## 1.1 Load essential packages
```{r, message = FALSE}
library(tidyverse)
library(openxlsx)
library(devtools)
install.packages("devtools")
library(icd)
```

## 1.2 Load and process data
```{r, message = FALSE}
bsi_mort <- read_csv("bsi_mort.csv")


.. cols(
  ..   pt_id = col_double(),
  ..   sex = col_character(),
  ..   dob...3 = col_character(),
  ..   age = col_double(),
  ..   Differenza_Prel_Ricov = col_double(),
  ..   ING_DEG = col_character(),
  ..   ING_DEG_2 = col_character(),
  ..   speciality = col_character(),
  ..   directorate = col_character(),
  ..   ...10 = col_logical(),
  ..   type = col_character(),
  ..   Pathogen = col_character(),
  ..   carba_resistant = col_double(),
  ..   count = col_double(),
  ..   admit_id = col_double(),
  ..   year = col_double(),
  ..   month = col_double(),
  ..   study_month = col_double(),
  ..   los = col_double(),
  ..   death = col_double(),
  ..   `ICD-10-CM Diagnosis Codes` = col_character(),
  ..   description = col_character(),
  ..   order = col_double(),
  ..   dob...24 = col_character(),
  ..   dataric = col_character(),
  ..   sample_date = col_character(),
  ..   dc_date = col_character(),
  ..   date_death = col_character(),
  ..   days_admit_sample = col_double(),
  ..   ...30 = col_double(),
  ..   diff = col_double()
  .. )

  ..   ING_DEG = col_character(),
  ..   ING_DEG_2 = col_character(),
  ..   Department = col_character(),
  ..   sample_type = col_character(),
  ..   Pathogen = col_character(),
  ..   contatore = col_double(),
  ..   IDadm = col_double(),
  ..   IdPaz = col_double(),
  ..   admcode = col_double(),
  ..   dataric = col_datetime(format = ""),
  ..   DataDecesso = col_datetime(format = ""),
  ..   DimessoDecesso = col_character(),
  ..   DataDim = col_datetime(format = "%Y-%m-%d  %H:%M:%S"),
  ..   DurataRicovero = col_double(),
  ..   `ICD-10-CM Diagnosis Codes...23` = col_character(),
  ..   description = col_character(),
  ..   order = col_double()
  .. )
```
