---
title: "Modena Time Series Analysis"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  pdf_document:
    toc: yes
---

# Introduction
This R notebook provides code to reproduce results for the paper *Effects of antimicrobial stewardship and infection prevention and control on the incidence of carbapenem resistant gram-negative infections in Modena, Italy: a non-linear time-series analysis*. 
It requires installing of necessary packages


Data and code are available at the Github repository
[via this link](https://github.com/TimLawes/modena_carbapenem_resistance)

# 1. Preliminaries
## 1.1 Install copies of required packages to private library
```{r, message=FALSE, warning=FALSE}
# Install renv for private package library if not available
if(!require(renv)){
    install.packages("renv")
    library(renv)
}
renv::restore()
```

## 1.2 Load required packages and customised plot theme
```{r, message=FALSE, warning=FALSE}
library(tidyverse) # to enable tidyverse functions
library(earth) # for MARS
```

## 1.3 Load data and process
```{r, message=FALSE, warning=FALSE}
main <- read_csv
```

## 1.4 Outlier detection and time-series cleaning
```{r, message=FALSE, warning=FALSE}

```
# 2. Descriptives

# 3. VARs

# 4. GAMs

# 5. MARS



```{r}
plot(cars)
```

