---
title: "Modena Time Series Analysis"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  pdf_document:
    toc: yes
---

# Introduction
This R notebook provides code to reproduce results for the paper *Effects of antimicrobial stewardship and infection prevention and control on the incidence of carbapenem resistant gram-negative infections in Modena, Italy: a non-linear time-series analysis*. 
It requires installing of necessary packages


Data and code are available at the Github repository
[via this link](https://github.com/TimLawes/modena_carbapenem_resistance)

# 1. Preliminaries
## 1.1 Install copies of required packages to private library
```{r message=FALSE, warning=FALSE, include=FALSE}

# Install renv for private package library if not available
if(!require(renv)){
    install.packages("renv")
}

Sys.setenv(LOCALAPPDATA=renv::paths$library())
renv::restore()
renv::activate()

```

## 1.2 Load required packages and customised plot theme
```{r, message=FALSE, warning=FALSE}

library(tidyverse) # to enable tidyverse functions
library(earth) # for MARS
library(tseries) # for working with TS data
library(tsutils) #
library(forecast) # for forcasting
library(fpp3) # enables tsibble, and fable packages
library(plotmo) # plotting MARS model from earth
library(mgcv) # for GLM and GAM models
library(smooth)
library(lmtest) # comparing models 
library(stats)
library(grid)
library(lattice)
library(viridis) # for viridis colour palettes
library(hrbrthemes) #for plot themes
library(sysfonts) # for using google fonts
library(lubridate) # for working with datetime variables
library(car) # for Cook's distances
library(strucchange) # for detecting structural changes in time-series
library(ganttrify) # for time lines (Gantt type)
library(patchwork) # for combining different plot types
library(missForest) # for missing value imputation by random forest
source("cochrane_plot_theme.R") # imports a custom ggplot theme defined in other script
source("OutlierTreatment2.R") # imports custom function for detecting extreme outliers and replacing.
```

## 1.3 Load data and process
```{r, message=FALSE, warning=FALSE, results='hide'}

# Read in the main data and convert to a time-series table (tsibble) with "tsmonth" as the index variable
main <- read_csv ("modena_carba_final.csv",
                  show_col_types = FALSE)

main <- main %>% mutate(month2 = as.POSIXct(paste("01-", Month, sep = ""), format = "%d-%b-%y"))

tsmain <- main %>% mutate(tsmonth = yearmonth(month2)) %>%
  as_tsibble(index = tsmonth)


# Calculate incidence of infections and colonisations (non-duplicate cases per 10 000 OBDs) and standardise antibiotic and alcohol-based hand rub use (DDDs or Litres per 1000 OBDs)
tswide <- tsmain  %>% mutate(across(
  gmneg_all_bsi:crpa_cohca_swab,~./obd*10000))

tswide <- tswide %>% mutate(across(abhr: fluoroquinolones, ~./obd*1000))
tsablong <- tswide %>% as_tibble() %>% dplyr::select(study_m, all_monitor_abx:fluoroquinolones) %>% pivot_longer(!study_m, names_to = "abx", values_to="use")

# create long format files
tslong <- tsmain %>% dplyr::select(!c(hh_adherence:mrsa_all_bsi)) %>%
  pivot_longer(
  cols = csab_all_any:crpa_cohca_swab,
  names_to = c("IR", "org", "pop", "site"),
  names_pattern = "(cr|cs)(ab|kp|pa)_(.*)_(.*)") %>% filter(pop!="NA") %>% dplyr::select(-c(Month, month2))

# create long format data for incidence
tslong_incid <- tslong %>% mutate(incidence = value/obd*10000)
tslong_incid <- tslong_incid %>% mutate (bugs = paste(IR,org,pop,site))

# create long format data for percent carbapenem resistance
tslong_pc <- tslong %>% pivot_wider(names_from="IR", values_from = "value") %>% mutate(pcent = cr/(cs+cr)*100)

tslong_pc$pcent <- dplyr::recode(tslong_pc$pcent,"0"="0.1")

# Read in intervention time-line
timeline <-read_csv("timelines.csv")

# Read in bloodstream infection dataset
blood <- read_csv("bsi_mort.csv")

# Read in additional data on changes in hand hygiene adherence by speciality
hhdata <- read_csv("hh_by.csv", col_types = cols (
  speciality = col_factor(),
  directorate = col_factor()))

# Impute missing data in period 2 of hand-hygiene by speciality dataset
hand <-as.data.frame(hhdata)
hhdata.imp <-missForest(hand)
adherence2 <-hhdata.imp[["ximp"]][["adherence"]] %>% as_tibble()
hhdata2 <- cbind(hhdata,adherence2)
rm(hhdata, adherence2)
rm(hhdata.imp)

```

## 1.4 Outlier detection and time-series cleaning
```{r, message=FALSE, warning=FALSE}

# outlier detection and replacement with cleaned values
tswideabx <- tswide %>% as_tibble %>% dplyr::select(obd,bedocc,
abhr:fluoroquinolones)

newabx.matrix<-rep(0,132)
for(i in 1:length(tswideabx[1,])){
abx<- tswideabx[,i]
newabx<-OutlierTreatment2(abx,freq=12,start.date=c(2008,1), only.clean=TRUE)
newabx.matrix <-cbind(newabx.matrix, newabx)
}
newabx.matrix<-newabx.matrix[,-1]

colnames(newabx.matrix)<-c(paste0(colnames(tswideabx)))
tswide_clean <-tswide %>% dplyr::select(!c(obd,bedocc,
abhr:fluoroquinolones))
tswide_clean <- cbind(tswide_clean, newabx.matrix) %>% as_tsibble(index = tsmonth)

tswide_clean %>% autoplot(gmneg_all_bsi)
tswide_cl_pre <- tswide_clean %>% filter(tswide_clean$int_hand == FALSE)
tswide_clean %>% autoplot(cephalosporins_DDD)

```


## 1.5 Trend decomposition
```{r, message=FALSE, warning=FALSE}
tslong <- tsmain %>% dplyr::select(!c(hh_adherence:mrsa_all_bsi)) %>%
  pivot_longer(
  cols = csab_all_any:crpa_cohca_swab,
  names_to = c("IR", "org", "pop", "site"),
  names_pattern = "(cr|cs)(ab|kp|pa)_(.*)_(.*)") %>% filter(pop!="NA") %>% dplyr::select(-c(Month, month2))

# create long format data for incidence
tslong_incid <- tslong %>% mutate(incidence = value/obd*10000)
tslong_incid <- tslong_incid %>% mutate (bugs = paste(IR,org,pop,site))

#####################
tslong_bugs <- tswide %>% dplyr::select(-(bedocc:mrsa_all_bsi)) %>%
  pivot_longer(
  cols = c(csab_all_any:crpa_cohca_swab),
  names_to = c("IR", "org", "pop", "site"),
  names_pattern = "(cr|cs)(ab|kp|pa)_(.*)_(.*)",
  values_to="incidence") %>% filter(incidence !="NA") %>% 
  select(-c("Month","study_m","month2","obd")) %>%
  mutate(bugs = paste(IR,org,pop,site))

tslong_bugs2 <- tslong_bugs %>% group_by(bugs,IR,org,pop,site) %>% nest() %>% filter(pop != "all")

decompose <- function(data, y, ...){
    y <- enquo(y)
    dvar <- data %>% pull(!!y) %>% as.ts()
    tmpd <- data %>% model(STL(dvar ~ trend(window = 7) + 
                                 season(window =13), robust = TRUE)) %>%
      components()
    data <-cbind(data,tmpd[,3:6])
}

ts_decomp <- tslong_bugs2 %>% 
    mutate(decs = map(data, decompose, y = incidence)) %>%
    dplyr::select(-data) %>% unnest(decs) %>% ungroup()
ts_decomp <- ts_decomp %>% mutate (trend=if_else(trend <0, 0, trend))

plotting of clinical infection type series
clinicals <- c("bsi","resp")
mainpops <- c("hca", "conothca")
ts_decomp_ab <- ts_decomp %>% filter (org == "ab") %>% filter(IR=="cr")

ts_decomp_crab <- ts_decomp_ab %>% filter (IR == "cr") %>% filter (site == "bsi" | site =="resp" | site == "uti" |site =="other") %>% filter (pop == "hca" | pop=="conothca")

 

ab_sites <- ggplot(ts_decomp_crab, aes(x=tsmonth, y=inc, group=site, color=site))+
  geom_line()+
  facet_grid (.~pop)
ab_sites+scale_colour_viridis_d(option="A", begin=0, end = 0.8)

```

# 2. Descriptives

## 2.1 Tab 1
```{r, message=FALSE, warning=FALSE}




```
## 2.2 Fig 1 CRO time series
```{r, message=FALSE, warning=FALSE}
# Create a generic timeline for the interventions
timing <- ganttrify(project = timeline,
          project_start_date = "2008-01",
          hide_wp=TRUE,
          size_text_relative = 0.9,
          colour_palette = inferno(4, begin=0.2, 
                                   end = 0.8),
          line_end = "square",
          month_breaks= 12,
          mark_years = TRUE,
          month_number_label = FALSE,
          month_date_label = TRUE,
          font_family = "Source Sans Pro ")

# Create plots for total incidence by pathogen and population


# create plots for numbers by carbapenem I/R and percentage I/R
mainpc_long <- mainpc %>% pivot_longer(
  cols= c(cs,cr), names_to="IR")
mainpc_long$site <- dplyr::recode(mainpc_long$site, admswab ="swab")

mainpc_long$IR <- dplyr::recode(mainpc_long$IR, cr ="I/R", cs= "S")
mainpc_long$IR <- as_factor(mainpc_long$IR)
mainpc_long$IR <- relevel(mainpc_long$IR, "S","I/R")
mainpop <- c("hca","conothca")
mainpc_long_ed <- mainpc_long %>% filter(pop %in% mainpop)

## plots for A.baumannii
ts_decomp_ab
ab_all <- mainpc_long_ed %>% filter (org == "ab")
ab_all <- ab_all %>% mutate(pcent = dplyr::recode(pcent, 
                                                  "0"= 0.1L))
infections <- c("bsi","resp","uti","other","swab")
ab_all2 <- ab_all %>% filter(site %in% infections)
ab_clinical <- mainpc_long_ed %>% filter(org =="ab" & site =="clinical")
ab_colors <-c('#fdbe85',"#fd8d3c")

abp1<- ggplot(ab_clinical, aes(x=study_m, y=value, fill=IR))+
  geom_col(width=1,alpha=0.7)+
  facet_grid(.~pop)+
  scale_y_continuous(name="% Carbapenem I/R")+
  scale_x_continuous(name = "Month and year", breaks=seq(1,132,12))+
  geom_vline(xintercept=57, color="grey",linetype="dotdash")+
  geom_vline(xintercept=73, color="grey",linetype="dotdash")+
  geom_vline(xintercept=81, color="grey",linetype="dotdash")+
  theme_pnaat()+
  theme(axis.text.x=element_blank())+
  scale_fill_manual (values=ab_colors)


abp2 <- ggplot(ab_all2, aes(x=study_m, y=site, fill=pcent))+
  geom_tile()+
  facet_grid(.~pop)+
  scale_x_continuous(name = "Month and year", breaks=seq(1,132,12))+
  scale_fill_viridis_c(option="B", begin  =0.2, end=1, na.value ="grey")


layoutab <- "
AAAA
AAAA
AAAA
BBBB
CCCC
CCCC
"
abp1 + abp2+ ab_sites+
  plot_layout(design = layoutab)
#### NEED TO ADD LINE TO ABOVE....suggest prior trend decomposition to allow long-term pattern to be seen ###

```

## 2.3 Fig 2 Antibiotic use

## 2.4 Fig 3 Hand hygiene improvements
```{r, message=FALSE, warning=FALSE}

# Time-series for Hand-hygiene adherence, training, and alcohol-based hand-rub use (Litres per 1000 OBDs)


t_breaks = seq(from = dmy('01-"jan"-08'), 
               to =dmy('01-"dec"-18'), by = '12 months')


hhp1 <- ggplot(maints, aes(month2, y= abhr))+
  geom_line(color="darkred")+
  scale_y_continuous(name="ABHR use*")+
  geom_vline(xintercept=as.POSIXct(as.Date(
     "2012-09-01")), linetype = "dotdash", 
     color="darkgrey")+
  scale_x_datetime(breaks= as.POSIXct(t_breaks),
                   date_labels="%b-%y")+
  theme_pnaat()+
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.title.y= element_text(size=9))
hhp1

hhp2 <- ggplot(maints, aes(month2, y= hh_adherence))+
  geom_line(color="darkred")+
  scale_y_continuous(name= "HH adherence, %",
                     limits=c(0,1), breaks=seq(0,1,0.2),
                     labels = function(y)
                       str_wrap(y, width = 5))+
  geom_vline(xintercept=as.POSIXct(as.Date(
     "2012-09-01")), linetype = "dotdash", 
     color="darkgrey")+
  scale_x_datetime(breaks= as.POSIXct(t_breaks),
                   date_labels="%b-%y")+
  theme_pnaat()+
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.title.y= element_text(size=9))
    

hhp3 <- ggplot(maints, aes(x=month2, 
                         y= hh_trained))+
  geom_line()+
  scale_y_continuous(name="Staff trained, n")+
  geom_vline(xintercept=as.POSIXct(as.Date(
     "2012-09-01")), linetype = "dotdash", 
     color="darkgrey")+
  scale_x_datetime(name = "Month and year", 
                   breaks= as.POSIXct(t_breaks),
                   date_labels="%b-%y")+
  theme_pnaat()+
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size=8),
        axis.title.y= element_text(size=9))+ scale_color_viridis_d(option="A")


# Hand-hygiene adherence by speciality and department
hhdata2$speciality <- fct_reorder2(
  hhdata2$speciality,hhdata2$period, hhdata2$adherence,
  first2, .desc=FALSE)

hhp4 <- ggplot(hhdata2, aes(
  x=period, y= speciality, fill=value))+
  geom_tile()+
  scale_y_discrete(name="Speciality")+
  scale_x_discrete(name="Intervention phase")+
  theme_pnaat()+
  scale_fill_viridis_c(option="A", begin=0, end=1)+
  labs(fill="Adherence")+
  theme(axis.title.x = element_text(size=9),
    axis.title.y= element_text(size=9),
    axis.text.y =element_text(size=8),
    legend.title = element_text(size=9),
    legend.text = element_text(size=8))

layout <- "
EEEED
AAAAD
BBBBD
CCCCD
"
hhp1 + hhp2 + hhp3 + hhp4 + timing+
  plot_layout(design = layout)

```
## 2.5 Fig 4 Screening

# 3. ARIMAs 
```{r}
# Look at key variables
tswide_clean %>% features(gmneg_all_bsi, unitroo_kpss)
tswide_clean  %>%
  features(gmneg_all_bsi, unitroot_kpss)
tswide_clean %>%
  mutate(diff_gmneg_all_bsi=difference(gmneg_all_bsi))

tswide_clean %>% features(gmneg_all_bsi, unitroot_nsdiffs, differences =0:2, .period=12)


tswide_clean %>%
  mutate(diff_gmneg_all_bsi=difference(gmneg_all_bsi)) %>%
  features(diff_gmneg_all_bsi, unitroot_ndiffs, differences =0:2, .period=1)

unitroot_nsdiffs()
%>%
  features(diff_gmneg_all_bsi, unitroot_kpss)

tswide_clean$diff_gmneg_all_bsi

# some example Cross-correlation functions 
## Between Gram-negative BSI and use of antibiotics
ccf(tswide_cl_pre$gmneg_all_bsi,tswide_cl_pre$cephalosporins_DDD)

## Between antibiotic use and CROs
### Klebsiella pneumoniae
ccf(tswide_cl_pre$carbapenems, tswide_cl_pre$crkp_hca_clinical)

auto.arima.df <- function(data, y, ...){
    y <- enquo(y)
    yts <- data %>% 
        pull(!!y) %>% 
        as.ts()
    auto.arima(yts, ...)
}
newabx.matrix<-rep(0,132)
for(i in 1:length(tswideabx[1,])){
abx<- tswideabx[,i]
newabx<-OutlierTreatment2(abx,freq=12,start.date=c(2008,1), only.clean=TRUE)
newabx.matrix <-cbind(newabx.matrix, newabx)
}
newabx.matrix<-newabx.matrix[,-1]

colnames(newabx.matrix)<-c(paste0(colnames(tswideabx)))
tswide_clean <-tswide %>% dplyr::select(!c(obd,bedocc,
abhr:fluoroquinolones))
tswide_clean <- cbind(tswide_clean, newabx.matrix) %>% as_tsibble(index = tsmonth)


```

# 4. Interactions and VARs

# 5. GAMs

# 6. MARS



```{r}
plot(cars)
```

