---
title: "Modena Time Series Analysis"
output:
  html_document:
    df_print: paged
    toc: yes
  word_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

# Introduction
This R notebook provides code to reproduce results for the paper *Effects of antimicrobial stewardship and infection prevention and control on the incidence of carbapenem resistant gram-negative infections in Modena, Italy: a non-linear time-series analysis*. 
It requires installing of necessary packages


Data and code are available at the Github repository
[via this link](https://github.com/TimLawes/modena_carbapenem_resistance)

# 1. Preliminaries
## 1.1 Install copies of required packages to private library
```{r  message=FALSE, warning=FALSE, results='hide'}

# Install renv for private package library if not available
if(!require(renv)){
    install.packages("renv")
}

Sys.setenv(LOCALAPPDATA=renv::paths$library())
renv::restore()
renv::activate()

```

## 1.2 Load required packages and customised plot theme
```{r message=FALSE, warning=FALSE, results='hide'}

library(tidyverse) # to enable tidyverse functions
library(scales) # additional functions for ggplot
library(labelled) # adding labels to variables (to appear in outputs)
library(earth) # for MARS
library(tseries) # for working with TS data
library(tsutils) #
library(forecast) # for forcasting
library(fpp3) # enables tsibble, and fable packages
library(timetk) # furtehr functions for wrangling and working with time-series
library(plotmo) # plotting MARS model from earth
library(mgcv) # for GLM and GAM models
library(itsadug) # additional functions for GAM and GAMM models
library(gratia) # plotting of GAMs and simultaneous CI estimation
library(smooth)
library(lmtest) # comparing models 
library(stats)
library(grid)
library(lattice)
library(viridis) # for viridis colour palettes
library(hrbrthemes) #for plot themes
library(sysfonts) # for using google fonts
library(imputeTS) # for imputing missing TS values
library(lubridate) # for working with datetime variables
library(car) # for Cook's distances
library(strucchange) # for detecting structural changes in time-series
library(ganttrify) # for time lines (Gantt type)
library(patchwork) # for combining different plot types
library(missForest) # for missing value imputation by random forest
library(mgcViz) # for plots and model checks of GAM
library(readxl)
source("cochrane_plot_theme.R") # imports a custom ggplot theme defined in other script

source("OutlierTreatment2.R") # imports custom function for detecting extreme outliers and replacing

source("smooth_diff_function.R") # Gavin Simpson's function for deriving smooth differences from an Xp matrix 

```



## 1.3 Load data and process
```{r warning=FALSE, include=FALSE, messages=FALSE, results='hide'}

#### 1) Read in and prepare "wide" format TS dataset ########

# Read in the main data 
main <- read_excel("../Data/raw data/modena_final_2008_2018.xlsx")

# Add date variable
main$date <- yearmonth(seq(as.Date("2008-01-01"),
                           as.Date("2018-12-31"), 
                           by = "1 month"))

# Add pc_variables


# Calculate incidences and create a time-series table (tsibble) with date as the index
tswide <- main %>% mutate(
  across(gmneg_all_bsi:crpa_cohca_swab,~./obds*10000),
  across(abhr: rectalswab, ~./obds*1000))  %>% 
  as_tsibble(index = date)

#  Select variables to evaluate for outliers (abx, obds, bed-occupancy, IPC, non-target GN pathogen incidences
wide_clean <- tswide %>% as_tibble %>% 
  dplyr::select(date, obds, bedocc,abhr:all_monitor_abx,gmneg_all_clinical:ecoli_esbl_all_bsi)

# Apply outlier detection /cleaning function and replace original variables
newabx.matrix<-rep(0,132) # creates empty matrix

for(i in 1:length(wide_clean[1,])){
abx<- wide_clean[,i]
newabx<-OutlierTreatment2(abx,freq=12,start.date=c(2008,1), only.clean=TRUE)
newabx.matrix <-cbind(newabx.matrix, newabx)
}

newabx.matrix<-newabx.matrix[,-1]

colnames(newabx.matrix)<-c(paste0(colnames(wide_clean)))
newabx.matrix<- newabx.matrix[,!colnames(newabx.matrix) %in% 'date']

tmp <-tswide %>% dplyr::select(!c(date, obds, bedocc,abhr:all_monitor_abx,gmneg_all_clinical:ecoli_esbl_all_bsi))

tswide_clean<-cbind(newabx.matrix,as.data.frame(tmp)) %>% as_tsibble(index=date)

rm(tmp, newab.matrix, date, abx, tswide, wide_clean) # Remove temporary data

# generate new vars after box-cox transformation
tswide_clean <- tswide_clean %>% mutate(across(
  c(abhr:ecoli_esbl_all_bsi, allab_all_clinical:crpa_cohca_swab, cephs,noncarb), box_cox_vec,.names="{.col}_bc"))

# label data
tswide_clean<- tswide_clean %>%
  set_variable_labels(
    bedocc= "Bed-occupancy (%)",
    abhr= "Alcohol-based hand rub use",
    rectalswab = "Rectal swabs",
    fq = "Fluoroquinolones",
    pseudo_3gc="Anti-pseudomonal cephalosporins",
    other_monitor_abx = "Other monitored antibiotics",
    all_monitor_abx = "All monitored antibiotics",
    phase = "Intervention phase",
    gmneg_all_clinical="Gram negative infections",
    ecoli_esbl_all_clinical="ESBL+ E.coli infections")




# create datasets for preintervention periods or sub-periods
tswide_clean$phase <-as_factor(tswide_clean$phase)
tswide_clean_p1 <- tswide_clean %>% filter(phase=="1")
tswide_clean_p2 <- tswide_clean %>% filter(phase=="2")
tswide_clean_p3 <- tswide_clean %>% filter(phase=="3")
tswide_clean_p4 <- tswide_clean %>% filter(phase=="4")
ts_i1 <- tswide_clean %>% filter(phase =="1")
ts_i2 <- tswide_clean %>% filter(phase =="1" | phase=="2")
ts_i3 <- tswide_clean %>% filter(!phase =="4")



#### 2) Generate "long" format dataset for microbial data ########
# for incidences
tsw <-tswide_clean %>% dplyr::select(!contains("_bc"))
tsw_bc <-tswide_clean %>% dplyr::select(contains("_bc"), date,phase)

tslong <- tsw %>% 
  dplyr::select(date, phase,
                   allab_all_clinical:crpa_cohca_swab) %>%
  pivot_longer(
  cols = allab_all_clinical:crpa_cohca_swab,
  names_to = c("IR", "org", "pop", "site"),
  names_pattern = "(cr|cs|all)(ab|kp|pa)_(.*)_(.*)",
  values_to="incidence") %>% filter(pop!="NA")

tslong_bc <- tsw_bc %>% 
  dplyr::select(date, phase,
                   allab_all_clinical_bc:crpa_cohca_swab_bc) %>%
  pivot_longer(
  cols = allab_all_clinical_bc:crpa_cohca_swab_bc,
  names_to = c("IR", "org", "pop", "site"),
  names_pattern = "(cr|cs|all)(ab|kp|pa)_(.*)_(.*_bc)",
  values_to="incidence") %>% filter(pop!="NA")

tslong<-dplyr::bind_rows(tslong, tslong_bc)

tslong<- tslong %>% mutate (bugs = paste(IR,org,pop,site))

tslong$phase <-as_factor(tslong$phase)
tslong_p1 <- tslong %>% filter(phase=="1")
tslong_p2 <- tslong %>% filter(phase=="2")
tslong_p3 <- tslong %>% filter(phase=="3")
tslong_p4 <- tslong %>% filter(phase=="4")



tslong<- tslong %>% mutate(bc=ifelse(grepl("_bc",site),1,0))

#Create nested long datasets
tslong_cro_nested <- tslong %>% group_by(bugs,IR,org,pop,site) %>% nest() 

```


# 2. Descriptives table by period

```{r message=FALSE, warning=FALSE, results='hide'}

# 3. Description of CRO and CSO time-series
sum_o <- tslong %>% features(incidence, 
                    feature_set(pkgs = "feasts")) %>% as_tibble()
sum_o_mean <- tslong %>% features(incidence, list(mean=mean,
                                  quantile)) %>% as_tibble()

sumstats_allstudy<- left_join(sum_o, sum_o_mean, by=c("IR","site","pop","org"))

sumstats_allstudy<- sumstats_allstudy  %>% mutate(
  bc=ifelse(grepl("_bc",site),1,0),
  outcome=paste0(IR,org,"_",pop,"_", site))

clins<-c("bsi","clinical","other","resp","uti")

ggplot(sumstats_allstudy %>% filter(
  site %in% clins & pop=="hca"), 
  aes(seasonal_peak_year, seasonal_strength_year, 
      color=site, size=mean))+
  geom_point(alpha=0.5)+facet_grid(org~IR)+
  scale_y_continuous(name="Strength of seasonality",
                     limits=c(0,0.7))+
  scale_x_continuous(name="Month of seasonal peak",
                     breaks=seq(0,12,3))+
  theme(panel.grid.minor.x=element_blank(),
        panel.grid.major.y=element_blank(),
        panel.grid.minor.y=element_blank())+
  scale_color_viridis_d(option="A", begin=0.2, end=0.8)


library(broom)
pcs <- sumstats_allstudy %>%
  dplyr::select(!c(IR, pop, site, org, mean:bc, nsdiffs)) %>%
  prcomp(scale = TRUE)
  
pcs %>%
  ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1)

```

# 3. Descriptives of trends

## 3.1 Change point analysis
```{r message=FALSE, warning=FALSE, results='hide'}
datcorr<- tswide_clean %>% as_tibble()  %>% 
  dplyr::select(crab_hca_clinical,crpa_hca_clinical, crkp_hca_clinical)

datab<- tswide_clean %>% as_tibble()  %>% 
  dplyr::select(crab_hca_clinical,crab_hca_bsi,
                crab_hca_uti, crab_hca_resp,crab_hca_other)

datpa<- tswide_clean %>% as_tibble()  %>% 
  dplyr::select(crpa_hca_clinical,crpa_hca_bsi,
                crpa_hca_uti, crpa_hca_resp,crpa_hca_other)

datkp<- tswide_clean %>% as_tibble()  %>% 
  dplyr::select(crkp_hca_clinical,crkp_hca_bsi,
                crkp_hca_uti, crkp_hca_resp,crkp_hca_other)

kcpRS_workflow(datcorr, RS_funs = c("runMean"), wsize = 3, K=10)
kcpRS_workflow(datkp, RS_funs = c("runMean"), wsize = 5, K=10)
kcpRS_workflow(datpa, RS_funs = c("runMean"), wsize = 5, K=10)
kcpRS_workflow(datab, RS_funs = c("runMean"), wsize = 5, K=10)  
kcp<- function(x, n){
List <- list()
for (i in n:1) {
l<- kcpRS_workflow(x, RS_funs = c("runMean"), wsize = n)
window <-rep(l$kcpMean$wsize, length(l$kcpMean$changePoints))
cp<- l$kcpMean$changePoints
lowcp<-round(cp-(window/2))
highcp<-round(cp+(window/2))
out<- cbind(window, cp, lowcp, highcp) %>% as.data.frame()
List[[length(List) + 1]]<-out
}
}


kcpRS(datkp, RS_fun=runMean, RS_name="Mean",Kmax=5, wsize=15)
kcpRS(datkp, RS_fun=runMean, RS_name="Mean", wsize = 25, nperm = 1000, Kmax = 10,
alpha = 0.05, varTest = FALSE, ncpu = 1)

look<- runMean(datcorr, wsize = 11)
look2<- runCorr(datcorr, wsize=11)
date<- seq(as.Date("2008-11-1"), as.Date("2018-12-1"), by = "months")
look<-cbind(date,look) 



looklong <- look %>% pivot_longer(cols=c(!date),
  names_to = c("IR", "org", "pop", "site"),
  names_pattern = "(cr|cs|all)(ab|kp|pa)_(.*)_(.*)",
  values_to="measure") %>% filter(pop!="NA")

looklong <- looklong %>% mutate(bugs=paste(IR,org,pop,site))


ggplot(looklong, aes(date, org, color=measure, fill=measure))+geom_tile()+
  scale_fill_viridis_c(option="A")+
  scale_color_viridis_c(option="A")

### R beast
# univariate
#CRKp
crkp<- tswide_clean %>% as_tibble()  %>% 
  dplyr::select(crkp_hca_clinical)
crkp<-ts(crkp, start=c(2008,1), frequency=12)
crkp_cp = beast123(crkp,metadata=metadata, season='none')

#CRPa
crpa<- tswide_clean %>% as_tibble()  %>% 
  dplyr::select(crpa_hca_clinical)
crpa<-ts(crpa, start=c(2008,1), frequency=12)
crpa_cp = beast123(crpa,metadata=metadata, season='none')

#CRAb
crab<- tswide_clean %>% as_tibble()  %>% 
  dplyr::select(crab_hca_clinical)
crab<-ts(crab, start=c(2008,1), frequency=12)
crab_cp = beast123(crab,metadata=metadata, season='none')

plot(crkp_cp)
plot(crpa_cp)
plot(crab_cp)
```

```{r message=FALSE, warning=FALSE, results='hide'}
# multivariate
datcorr<- tswide_clean %>%
  dplyr::select(date, crab_hca_clinical,crpa_hca_clinical, crkp_hca_clinical)


metadata<- list(freq=12, deltaTime=1, 
                       whichDimIsTime=2,
                       startTime=c(2008,1),
                       isRegularOrdered=T)
prior<- list(seasonMinKnotNum=1, seasonMaxKnotNum=3,
                    trendMinOrder=1, trendMaxOrder=4,
                     trendMinOrder=1, trendMaxOrder=3,
                    trendMinSepDist=6,
                    trendMinKnotNum=1, trendMaxKnotNum=3)
extra<- list(dumpInputData=T,whichOutputDimIsTime=2)

o = beast123(datcorr,metadata=metadata,prior=prior, extra=extra, season='dummy')

o = beast123(Yellowstone,metadata,prior,extra, season='harmonic')
plot(o)

crkp<- tswide_clean %>% as_tibble()  %>% 
  dplyr::select(crkp_hca_clinical)

crkp<-ts(crkp, start=c(2008,1), frequency=12)
```

## 3.2 Comparison of seasonality
```{r message=FALSE, warning=FALSE, results='hide'}

## Look at seasonality
seasonlook <- tslong %>% dplyr::select(bugs, incidence) %>%
  features(incidence, feat_stl)
incidences <- tslong %>% as_tibble() %>% group_by(bugs, IR, pop, site, org) %>% summarise("Mean monthly incidence rate" =mean(incidence))

seasonlook<-left_join (seasonlook, incidences, by=c("IR","pop","site","org"))

seasonlook$IR <- fct_recode(seasonlook$IR, 
                            "Carbapenem sensitive"="cs", 
                            "Carbapenem resistant"="cr")
seasonlook$org <- fct_recode(seasonlook$org, 
                            "A.baumannii"="ab", "K.pneumoniae"="kp", 
                            "P.aeruginosa"="pa")
seasonlook$pop <- fct_recode(seasonlook$pop, 
                            "HCA"="hca", "CO-HCA"="cohca", "All"="all",
                            "HO-HCA"="ho", "Non-HCA"="conothca")
seasonlook$site <- fct_recode(seasonlook$site, "all"="any")


seasonlook$peak =seasonlook$seasonal_peak_year+1
require(lubridate)
seasonlook <- seasonlook %>% mutate(month=parse_date_time(peak, "%m"))

seasonality<- ggplot(seasonlook %>% filter(site=="clinical"),
       aes(x=month, 
           y = seasonal_strength_year, 
           size=`Mean monthly incidence rate` , 
           col = pop, fill=pop)) +
  geom_point(alpha=0.4) +
  facet_grid(org~IR)+
  scale_color_viridis_d(option="A", begin=0.2,end=0.8)+
  scale_x_datetime("Month of seasonal peak", 
                   labels=date_format("%b"))+
  scale_y_continuous("Strength of seasonality", limits=c(0,0.75))+
  theme_pnaat()

seasonality

seasonality_inf_cr<- ggplot(
  seasonlook %>% filter(site=="clinical" & IR=="Carbapenem resistant"),
       aes(x=month, 
           y = seasonal_strength_year, 
           size=`Mean monthly incidence rate` , 
           col = pop, fill=pop)) +
  geom_point(alpha=0.4) +
  facet_grid(org~.)+
  scale_color_viridis_d(option="A", begin=0.2,end=0.8)+
  scale_x_datetime("Month of seasonal peak", 
                   labels=date_format("%b"))+
  scale_y_continuous("Strength of seasonality", limits=c(0,0.75))+
  theme_pnaat()

seasonality_inf_cr

seasonality_inf_cs<- ggplot(
  seasonlook %>% filter(site=="clinical" & IR=="Carbapenem sensitive"),
       aes(x=month, 
           y = seasonal_strength_year, 
           size=`Mean monthly incidence rate` , 
           col = pop, fill=pop)) +
  geom_point(alpha=0.4) +
  facet_grid(org~.)+
  scale_color_viridis_d(option="A", begin=0.2,end=0.8)+
  scale_x_datetime("Month of seasonal peak", 
                   labels=date_format("%b"))+
  scale_y_continuous("Strength of seasonality", limits=c(0,0.75))+
  theme_pnaat()

seasonality_inf_cs
```


## 3.3 Plot of TS 
```{r message=FALSE, warning=FALSE, results='hide'}

decompose <- function(data, y, ...){
    y <- enquo(y)
    dvar <- data %>% dplyr::pull(!!y) %>% as.ts()
    tmpd <- data %>% model(STL(dvar ~ trend(window = 13) + 
                                 season(window =13), 
                               robust = TRUE)) %>%
      components()
    data <-dplyr::bind_cols(data,tmpd[,3:6]) %>% as_tibble()
}

# TS decomposition for incidence
ts_decomp <- tslong_cro_nested %>% 
    mutate(decs = map(data, decompose, y = incidence)) %>%
    dplyr::select(-data) %>% unnest(decs)

ts_decomp$inf <- if_else(ts_decomp$site=="swab","swab","clinical")
ts_decomp <-ts_decomp %>% mutate(
  time = ym(as.character(date)))
ts_decomp$IR <-as_factor(ts_decomp$IR)
ts_decomp$org <-as_factor(ts_decomp$org)

ts_decomp$IR <- fct_recode(
  ts_decomp$IR, "Carbapenem sensitive"="cs", "Carbapenem resistant"="cr")
ts_decomp$org <- fct_recode(
  ts_decomp$org, "A.baumannii"="ab", "K.pneumoniae"="kp","P.aeruginosa"="pa")

ts_decomplong<- ts_decomp %>% pivot_longer(cols=c("trend","season_year","remainder"),names_to="component")

# TS decomposition for percentages
ts_decomp_pc <- tslong_pc_cro_nested %>% 
    mutate(decs = map(data, decompose, y = pc_cro)) %>%
    dplyr::select(-data) %>% unnest(decs)




cro_mainplot<- ggplot(ts_decomp %>% 
    filter(pop=="hca" & site=="clinical" & IR=="Carbapenem resistant"), 
       aes(time, trend, color=org, ))+
  geom_line()+
  geom_line(aes(x=time, y=incidence), alpha=0.3)+
  facet_grid(org~., scales="free_y")+
  scale_color_viridis_d(option="A",begin=0.2, end=0.8)+
  theme_pnaat()+
  theme(panel.grid=element_blank())+
  scale_x_date("Year", date_breaks="1 year", date_labels = "%Y")+
  scale_y_continuous("All clinical infections (cases per 10 000 OBDs)")+
  geom_vline(xintercept=as.numeric(ts_decomp$time[57]),
             linetype=2, color="grey")+
  geom_vline(xintercept=as.numeric(ts_decomp$time[73]),
             linetype=2, color="grey")+
  geom_vline(xintercept=as.numeric(ts_decomp$time[81]),
               linetype=2, color="grey")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        legend.position="none")

cso_mainplot<- ggplot(ts_decomp %>% 
    filter(pop=="hca" & site=="clinical" & IR=="Carbapenem sensitive"), 
       aes(time, trend, color=org, ))+
  geom_line()+
  geom_line(aes(x=time, y=incidence), alpha=0.3)+
  facet_grid(org~., scales="free_y")+
  scale_color_viridis_d(option="A",begin=0.2, end=0.8)+
  theme_pnaat()+
  theme(panel.grid=element_blank())+
  scale_x_date("Year", date_breaks="1 year", date_labels = "%Y")+
  scale_y_continuous("All clinical infections (cases per 10 000 OBDs)", limits=c(0,NA))+
  geom_vline(xintercept=as.numeric(ts_decomp$time[57]),
             linetype=2, color="grey")+
  geom_vline(xintercept=as.numeric(ts_decomp$time[73]),
             linetype=2, color="grey")+
  geom_vline(xintercept=as.numeric(ts_decomp$time[81]),
               linetype=2, color="grey")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        legend.position="none")


pc_plot<- ggplot(
  ts_decomp %>% filter(pop=="hca" & site=="clinical"), 
       aes(time, trend, color=org, ))+
  geom_line()+
  geom_line(aes(x=time, y=incidence), alpha=0.3)+
  facet_grid(org~IR, scales="free_y")+
  scale_color_viridis_d(option="A",begin=0.2, end=0.8)+
  theme_pnaat()+
  theme(panel.grid=element_blank())+
  scale_x_date("Year", date_breaks="1 year", date_labels = "%Y")+
  scale_y_continuous("All clinical infections (cases per 10 000 OBDs)", limits=c(0,NA))+
  geom_vline(xintercept=as.numeric(ts_decomp$time[57]),
             linetype=2, color="grey")+
  geom_vline(xintercept=as.numeric(ts_decomp$time[73]),
             linetype=2, color="grey")+
  geom_vline(xintercept=as.numeric(ts_decomp$time[81]),
               linetype=2, color="grey")

layout <- "AAAAABBBBBC"

cro_mainplot+ cso_mainplot

```


#  4. Regional & national trends

## 2.1 preparation of data
```{r  message=FALSE, warning=FALSE,  results='hide', cache= TRUE}

# Prepare data
otherpops <- read_csv("../Data/raw data/populations_year.csv", col_types = cols(
  pop = col_factor(),
  org = col_factor(),
  year = col_integer(),
  site = col_factor(),
  num_ts=col_factor()))


otherpops <-otherpops %>% filter(studyp==1 & carb_R != "NA") %>%
  mutate(pop1 = pop)
otherpops$pop1 <- as_factor(otherpops$pop1) 
otherpops$pop1 <- fct_recode(otherpops$pop1, M="Modena", ER = "Emilia-Romagna", ERo = "Emilia-Romagna (other)", I = "Italy")
otherpops <- otherpops %>% mutate(
    poporg = interaction(pop1, org), 
         sitepop = interaction(site,pop1))
otherpops$poporg <-as_factor(otherpops$poporg)
otherpops$sitepop <-as_factor(otherpops$sitepop)

#### Data for evaluating number tested by population
otherpops_num <- otherpops %>% filter(pop!="Emilia-Romagna") %>%
  dplyr::select(year:tested) %>% pivot_longer (cols= c(carb_S, carb_I, carb_R),
                     names_to = "Susceptibility", values_to = "Number")

otherpops_num$Susceptibility<- as_factor(otherpops_num$Susceptibility)%>% fct_recode(R = "carb_R", I = "carb_I", S="carb_S")
otherpops_num$pop<- fct_recode(otherpops_num$pop, "ER other" =  "Emilia-Romagna (other)")



#### Data for regional isolate results

otherpops_pc <- otherpops %>% filter(
  pop=="Modena"| pop=="Emilia-Romagna") %>%
  dplyr::select(year:carb_SI) %>% 
  rename(SI = "carb_SI", R = "carb_R")

otherpops_pc <- otherpops_pc %>% pivot_wider(id_cols=c(year,org,site),
                names_from = pop, values_from=c(R, SI))

otherpops_pc <- otherpops_pc %>% pivot_longer(
  cols=c("R_Modena","SI_Modena", "R_Emilia-Romagna",
         "SI_Emilia-Romagna"),
  names_to=c("Susceptibility",".value"), values_to = "popn",
  names_pattern=("(.*)_(.*)"))

otherpops_pc$Susceptibility <- as_factor(otherpops_pc$Susceptibility)

otherpops_pc <- otherpops_pc %>% mutate(
  total = Modena+`Emilia-Romagna`, 
  pc = Modena / total, 
  siteorg = interaction(site,org))

```


## 2.2 Number tested and Percentage of all isolates within 
```{r  message=FALSE, warning=FALSE,  results='hide', cache= TRUE}


#### GAM model of % IR or S All clinical isolates in region from Modena
gdclinical <- otherpops_pc %>% filter(site == "All clinical")

pcregion_GAM <- gam(Modena/`Emilia-Romagna`~ org*Susceptibility+ s(year, by=interaction(Susceptibility,org),
                    m=1, k=-1, bs="tp"), 
                    data = gdclinical, weights = total, 
                    family = binomial(link="logit"), method="REML")

# get model summary
pcregionViz <- getViz(pcregion_GAM)
summary.gam(pcregion_GAM)  
BIC(pcregionViz)


# standard model fit diagnostics
appraise(pcregion_GAM, method= "simulate", n_simulate = 1000, type = "deviance", level=0.9, point_col = "steelblue", point_alpha = 0.4,  ci_alpha = 0.1)

# check for adequacy of basis dimension.
k.check(pcregion_GAM)

# checks for autocorrelation
acf(resid(pcregion_GAM), plot=FALSE)$acf[2]

# checks for concurvity
concurvity(pcregionViz, full=TRUE)

# collate model fits 
pcregion_fit<- cbind(gdclinical, predict.gam(pcregion_GAM,
                  se.fit=TRUE, type = "link")) 
pcregion_fit$uci<-pcregion_fit$fit + (2 * pcregion_fit$se.fit) 
pcregion_fit$lci<-pcregion_fit$fit - (2 * pcregion_fit$se.fit) 
pcregion_fit$uci<-pcregion_GAM$family$linkinv(pcregion_fit$uci )
pcregion_fit$lci<-pcregion_GAM$family$linkinv(pcregion_fit$lci )
pcregion_fit$est<- pcregion_GAM$family$linkinv(pcregion_fit$fit)

pcregion_fit <- pcregion_fit %>% mutate(obs_pc = Modena/total)



#### GAM model of % IR or S BSI in region from Modena
# Filter data
gdbsi <- otherpops_pc %>% filter(site == "BSI")

pcregbsi_GAM <- gam(Modena/`Emilia-Romagna`~ org*Susceptibility+ s(year, by=interaction(Susceptibility,org),m=2, k=-1, bs="tp"), data = gdbsi, weights = total, family = binomial(link="logit"), method="REML")

# get model summary
pcregbsiViz <- getViz(pcregbsi_GAM)
summary.gam(pcregbsi_GAM)  
BIC(pcregbsiViz)

# standard model fit diagnostics
appraise(pcregbsi_GAM, method= "simulate", n_simulate = 1000, type = "deviance", level=0.9, point_col = "steelblue", point_alpha = 0.4,  ci_alpha = 0.1)

# check for adequacy of basis dimension.
k.check(pcregbsi_GAM)

# checks for autocorrelation
acf(resid(pcregbsi_GAM), plot=FALSE)$acf[2]

# checks for concurvity
concurvity(pcregbsiViz, full=TRUE)

# collate model fits 
pcregbsi_fit<- cbind(gdbsi, predict.gam(pcregbsi_GAM,
                  se.fit=TRUE, type = "link")) 
pcregbsi_fit$uci<-pcregbsi_fit$fit + (2 * pcregbsi_fit$se.fit) 
pcregbsi_fit$lci<-pcregbsi_fit$fit - (2 * pcregbsi_fit$se.fit) 
pcregbsi_fit$uci<-pcregbsi_GAM$family$linkinv(pcregbsi_fit$uci )
pcregbsi_fit$lci<-pcregbsi_GAM$family$linkinv(pcregbsi_fit$lci )
pcregbsi_fit$est<- pcregbsi_GAM$family$linkinv(pcregbsi_fit$fit)

pcregbsi_fit <- pcregbsi_fit %>% mutate(obs_pc = Modena/total)
pcregionall <- rbind(pcregbsi_fit, pcregion_fit)


# Final plot for numbers reported
ggplot(otherpops_num, 
                      aes(year, Number, fill=Susceptibility))+
  geom_col(colour=NA, alpha= 0.6)+
  facet_grid(pop+site~org, scales="free_y")+
  scale_y_continuous(name="Number of isolates tested")+
  scale_x_continuous(name = "Year", breaks=seq(2008,2018,2))+
  theme_pnaat()+
  scale_color_viridis_d(option="A", begin = 0.8, end=0.3)+
  scale_fill_viridis_d(option="A", begin = 0.8, end=0.3)+
  theme(axis.text.x = element_text(angle = 45, hjust =1))+
  theme(legend.position = "top")+
  theme(legend.title=element_text(size=9))

ggsave("./Output/regnat_num.jpg")


# Plot for proportion of regional isolates occurring in Modena by susceptibility, organism, body site
ggplot(pcregionall, aes(year, est, 
  group= Susceptibility, fill=Susceptibility, color= Susceptibility))+
  geom_line()+
  geom_point(aes(year, pc))+
  geom_ribbon(aes(x=year, ymin=lci, ymax=uci), color=NA, alpha=0.2)+
  scale_x_continuous("Year", breaks=seq(2008,2018,2))+
  facet_grid(site~org, scales="free_y")+
  scale_y_continuous("% Regional isolates within S/IR group",
                     labels=scales::percent_format(accuracy=1))+
  scale_color_viridis_d(option="A", begin=0.3, end=0.8)+
  scale_fill_viridis_d(option= "A", begin=0.3, end = 0.8)+
  theme_pnaat()+
  coord_cartesian(ylim=c(0, 0.3))+
  theme(axis.text.x = element_text(angle = 45, hjust =1))+
  theme(legend.position = "top")+
  theme(legend.title=element_text(size=9))


ggsave("./Output/regnat_pcmodena.jpg", 
       height=15, width=20, units="cm")


```


## 2.3 GAMs by organism
```{r message=FALSE, warning=FALSE, results='hide', cache= TRUE}

####  K.pneumoniae ########################################

## Filter data
kp_pop <- otherpops %>% filter(org =="K.pneumoniae")

## Model Smooth by Interaction of factors
kp_pop_GAMx <- bam(carb_R/tested ~ s(sitepop,bs="re") + s(year,by=sitepop,m=2, k=-1, bs="tp"), data =  kp_pop, weights = tested, family = binomial(link="logit"), method="REML", rho=0.5, AR.start=kp_pop$start.event)

# get model summary
kppopVizx <- getViz(kp_pop_GAMx)
summary.gam(kp_pop_GAMx)
BIC(kppopVizx)

# standard model fit diagnostics
appraise(kp_pop_GAMx, method= "simulate", n_simulate = 1000, type = "deviance", level=0.9, point_col = "steelblue", point_alpha = 0.4,  ci_alpha = 0.1)

# check for adequacy of basis dimension.
k.check(kp_pop_GAMx)

# checks for autocorrelation
acf(resid(kp_pop_GAMx), plot=FALSE)$acf[2]

# checks for concurvity
concurvity(kppopVizx, full=TRUE)

##### P.aeruginosa GAM ###################################

## Filter data
pa_pop <- otherpops %>% filter(org =="P.aeruginosa")

# Model Smooth by Interaction of factors
pa_pop_GAMx <- bam(carb_R/tested ~ s(sitepop,bs="re") + s(year,by=sitepop,m=2, k=-1, bs="tp"), data =  pa_pop, weights = tested, family = binomial(link="logit"), method="REML", rho=0.5, AR.start=pa_pop$start.event)

# get model summary
papopVizx <- getViz(pa_pop_GAMx)
summary.gam(pa_pop_GAMx)  
BIC(papopVizx)

# standard model fit diagnostics
appraise(pa_pop_GAMx, method= "simulate", n_simulate = 1000, type = "deviance", level=0.9, point_col = "steelblue", point_alpha = 0.4,  ci_alpha = 0.1)

# check for adequacy of basis dimension.
k.check(pa_pop_GAMx)  

# checks for autocorrelation
acf(resid(pa_pop_GAMx), plot=FALSE)$acf[2]

# checks for concurvity
concurvity(papopVizx, full=TRUE)


##### A.baumannii GAM
# data
ab_pop <- otherpops %>% filter(org =="A.baumannii")

# model
ab_pop_GAMx <- gam(carb_R/tested ~ s(sitepop,bs="re") + s(year,by=sitepop,m=2, k=-1, bs="tp"), data =  ab_pop, weights = tested, family = binomial(link="logit"), method="REML")

# get model summary
abpopVizx <- getViz(ab_pop_GAMx)
summary.gam(ab_pop_GAMx) 
BIC(abpopVizx)

# standard model fit diagnostics
appraise(ab_pop_GAMx, method= "simulate", n_simulate = 1000, type = "deviance", level=0.9, point_col = "steelblue", point_alpha = 0.4,  ci_alpha = 0.1)

# check for adequacy of basis dimension.
k.check(ab_pop_GAMx)

# checks for autocorrelation
acf(resid(ab_pop_GAMx), plot=FALSE)$acf[2]

# checks for concurvity
concurvity(abpopVizx, full=TRUE)


```

## 2.4 Plot GAM predictions

```{r message=FALSE, warning=FALSE, results='hide', cache= TRUE}

names <- c("All clinical.M",  "BSI.M", 
              "All clinical.ERo", 
              "BSI.ERo", "BSI.I", 
              "All clinical.ER", "BSI.ER")
            
#### Create data for fitting model predictions
pdat <- expand.grid(
  year= seq(2008, 2018, length = 264),
  sitepop = names)


#### Fit model predictions 
kp_pred <- predict(kp_pop_GAMx, pdat, type="response", se.fit=TRUE)
kp_pred <- cbind(kp_pred, pdat) %>% as_tibble() %>%
  separate(sitepop, into=c("site1","pop2"), sep="[.]") %>%
  mutate(org = "K.pneumoniae")

pa_pred <- predict(pa_pop_GAMx, pdat, type="response", se.fit=TRUE)
pa_pred <- cbind(pa_pred, pdat) %>% as_tibble() %>%
  separate(sitepop, into=c("site1","pop2"), sep="[.]") %>%
  mutate(org = "P.aeruginosa")

ab_pred <- predict(ab_pop_GAMx, pdat, type="response", se.fit=TRUE)
ab_pred <- cbind(ab_pred, pdat) %>% as_tibble()  %>%
  separate(sitepop, into=c("site1","pop2"), sep="[.]") %>%
  mutate(org = "A.baumannii")

orgpreds <- rbind(kp_pred, pa_pred, ab_pred)
rm(kp_pred, pa_pred, ab_pred)

##### Prepare data for plots
orgpredsed <- orgpreds %>% filter (pop2 !="ER")

orgpredsed <- orgpredsed %>% mutate(
  italyfit = fit, 
  italyse = se.fit,
  fit = ifelse(pop2 =="I" & year <2013 & org =="A.baumannii", 
               NA, fit),
  se.fit = ifelse(pop2 == "I" & year < 2013 & org == "A.baumannii",
              NA, se.fit),
  italyfit = ifelse(pop2 == "I" & year < 2013 & org == "A.baumannii",
                    italyfit, NA),
  italyse = ifelse(pop2 == "I" & year < 2013
                         & org == "A.baumannii", italyse, NA))

orgpredsed <- orgpredsed %>% mutate(
  italyse =ifelse(italyse >1 | italyse <0, NA, italyse))

orgpredsed <- orgpredsed %>% rename(Population = "pop2")
orgpredsed$Population <- fct_recode(orgpredsed$Population, Modena="M", "Emilia-Romagna (other)"="ERo", "Emilia-Romagna"="ER", Italy="I")

otherpops_ed3 <-otherpops %>% 
  rename(Population= "pop", site1 = "site")%>% 
  filter (Population != "Emilia-Romagna")


# Plot of observed values and GAM predictions 
ggplot(orgpredsed, aes(year, fit, color=Population, fill=Population))+
  geom_line()+
  geom_line(aes(year,italyfit), linetype = "dashed")+
  geom_ribbon(aes(year, ymin=italyfit-(1.96*italyse),
              ymax=italyfit+(1.96*italyse)), fill=NA, linetype="dotted")+
  geom_point(data= otherpops_ed3, 
             aes(x=year, y=pc_R))+
  geom_ribbon(aes(year, ymin=fit-(1.96*se.fit),
      ymax=fit+(1.96*se.fit)), alpha=0.2, color=NA)+
  facet_grid(org~site1, scales="free_y")+
  scale_x_continuous ("Year", breaks=seq(2008,2018,2))+
  scale_y_continuous(name="% Carbapenem resistant",
             labels=scales::percent_format(accuracy=1),
             limits=c(0,NA))+
  scale_color_viridis_d(option="A", begin=0.8, end=0.2)+
  scale_fill_viridis_d(option="A", begin = 0.8, end=0.2)+
  theme_pnaat()+
  theme(legend.title=element_text(size=9))+
  theme(axis.text.x = element_text(angle = 45, hjust =1))+
  theme(legend.position = "top")

ggsave("./Output/regnat_pcR_site_pop.jpg", 
       height=15, width=20, units="cm")

```


## 2.5 Differences in resistance trends between populations
```{r message=FALSE, warning=FALSE, results='hide', cache= TRUE}

#### Between populations

# K.pneumoniae
MvER_all <- smooth_diff(kp_pop_GAMx, pdat, "All clinical.M", "All clinical.ERo", "sitepop") %>% mutate(re1 = -0.5451407, re2= -0.4564834)

MvI_bsi <- smooth_diff(kp_pop_GAMx, pdat, "BSI.M", "BSI.I", "sitepop") %>% mutate(re1 = 0.3637846, re2= 0.7332048)

MvER_bsi <- smooth_diff(kp_pop_GAMx, pdat, "BSI.M", "BSI.ERo", "sitepop")%>% mutate(re1 = 0.3637846, re2= 0.1554677)

ERovsI_BSI <-smooth_diff(kp_pop_GAMx, pdat, "BSI.ERo", "BSI.I", "sitepop")%>%mutate(re1 = 0.1554677, re2= 0.7332048)

ERovsER_all<-smooth_diff(kp_pop_GAMx, pdat, "All clinical.ERo", "All clinical.ER", "sitepop")%>% mutate(re1 = -0.4564834, re2= -0.442675)

ERovsER_BSI <-smooth_diff(kp_pop_GAMx, pdat, "BSI.ERo", "BSI.ER", "sitepop") %>% mutate(re1 = 0.1554677, re2= 0.1918419)

compkp <- cbind(date = seq(2008, 2018, length = 264),
              rbind(MvER_all, MvI_bsi, MvER_bsi,ERovsI_BSI,
                    ERovsER_BSI, ERovsER_all))

compkp <- compkp %>% separate(col=pair, into=c("site1","pop1", "pop2"),
                              sep = "[.]", remove = TRUE) %>%
                      separate(col="pop1", into=c("pop1", "null"), 
                               sep = "[-]", remove = TRUE) %>%
                      dplyr::select(-"null") %>% mutate(org =
                              "K.pneumoniae", reorg = -2.039596)

# P.aeruginosa 
MvER_all <- smooth_diff(pa_pop_GAMx, pdat, "All clinical.M", "All clinical.ERo", "sitepop") %>% mutate(re1 = -0.3030567, re2= -0.03286934)

MvI_bsi <- smooth_diff(pa_pop_GAMx, pdat, "BSI.M", "BSI.I", "sitepop")%>% mutate(re1 = 0.05955559, re2= -0.0465038)

MvER_bsi <- smooth_diff(pa_pop_GAMx, pdat, "BSI.M", "BSI.ERo", "sitepop") %>% mutate(re1 = 0.05955559, re2= -0.06386308)

ERovsI_BSI <-smooth_diff(pa_pop_GAMx, pdat, "BSI.ERo", "BSI.I", "sitepop") %>% mutate(re1 = -0.06386308, re2= 0.4333925)

ERovsER_all<-smooth_diff(pa_pop_GAMx, pdat, "All clinical.ERo", "All clinical.ER", "sitepop") %>% mutate(re1 = -0.03286934, re2= -0.0465038)

ERovsER_BSI <-smooth_diff(pa_pop_GAMx, pdat, "BSI.ERo", "BSI.ER", "sitepop") %>% mutate(re1 = -0.06386308, re2= -0.04665511)

comppa <- cbind(date = seq(2008, 2018, length = 264),
              rbind(MvER_all, MvI_bsi, MvER_bsi,ERovsI_BSI,
                    ERovsER_BSI, ERovsER_all))

comppa <- comppa %>% separate(col=pair, into=c("site1","pop1", "pop2"),
                              sep = "[.]", remove = TRUE) %>%
                      separate(col="pop1", into=c("pop1", "null"), 
                               sep = "[-]", remove = TRUE) %>% dplyr::select(-"null") %>% mutate(org = "P.aeruginosa", reorg = -1.378)


# A.baumannii
MvER_all <- smooth_diff(ab_pop_GAMx, pdat, "All clinical.M", "All clinical.ERo", "sitepop") %>% mutate(re1 = 0, re2= 0)


MvI_bsi <- smooth_diff(ab_pop_GAMx, pdat, "BSI.M", "BSI.I", "sitepop") %>% mutate(re1 = 0, re2= 0)

MvER_bsi <-smooth_diff(ab_pop_GAMx, pdat, "BSI.M", "BSI.ERo", "sitepop") %>% mutate(re1 = 0, re2= 0)

ERovsI_BSI <-smooth_diff(ab_pop_GAMx, pdat, "BSI.ERo", "BSI.I", "sitepop") %>% mutate(re1 = 0, re2= 0)

ERovsER_all<-smooth_diff(ab_pop_GAMx, pdat, "All clinical.ERo", "All clinical.ER", "sitepop") %>% mutate(re1 = 0, re2= 0)

ERovsER_BSI <-smooth_diff(ab_pop_GAMx, pdat, "BSI.ERo", "BSI.ER", "sitepop") %>% mutate(re1 = 0, re2= 0)

compab <- cbind(date = seq(2008, 2018, length = 264),
              rbind(MvER_all, MvI_bsi, MvER_bsi,ERovsI_BSI,
                    ERovsER_BSI, ERovsER_all))

compab <- compab %>% separate(col=pair, into=c("site1","pop1",
                "pop2"), sep = "[.]", remove = TRUE) %>%
                separate(col="pop1", into=c("pop1", "null"), 
                sep = "[-]", remove = TRUE) %>%
                dplyr::select(-"null") %>% mutate(
                  org = "A.baumannii", reorg= 1.299869)

comppops <- rbind(compkp, comppa, compab)


comppops$pop1 <- as_factor(comppops$pop1) %>% fct_recode(
"ERo" = "Emilia")


comppops <- comppops %>% mutate(Comparison = paste(pop1, "vs", pop2),
            tdiff = diff+(re1-re2),
            tlci = lower+(re1-re2),
            tuci = upper+(re1-re2))

comppops$Comparison <- as_factor(comppops$Comparison)
comppops$Comparison <- fct_recode(comppops$Comparison, "Modena vs. Emilia-Romagna (other)"="M vs ERo", "Modena vs. Italy" = "M vs I", "Emilia-Romagna (other) vs. Italy"= "ERo vs I")

comppops <- comppops %>% rename(year = "date")

comppops <- left_join (comppops, orgpreds, by = c("org","pop2","site1","year")) 

comppops <- comppops %>% mutate(
  totdiff = plogis(tdiff+logit(fit))-fit,
  totlci = plogis(tlci+logit(fit))-fit,
  totuci = plogis(tuci+logit(fit))-fit)

comppops <- comppops %>% mutate(
  italytotdiff = totdiff, 
  italylci = totlci,
  italyuci = totuci,
  totdiff = ifelse(pop2 =="I" & year <2012 & org =="A.baumannii", 
               NA, totdiff),
  totlci = ifelse(pop2 == "I" & year < 2012 & org == "A.baumannii",
              NA, totlci),
  totuci = ifelse(pop2 == "I" & year < 2012 & org == "A.baumannii",
              NA, totuci),
  italytotdiff = ifelse(pop2 == "I" & year < 2012 & org ==
                          "A.baumannii",italytotdiff, NA),
  italylci = ifelse(pop2 == "I" & year < 2012 & org ==
                         "A.baumannii", italylci, NA),
  italyuci = ifelse(pop2 == "I" & year < 2012 & org ==
                         "A.baumannii", italyuci, NA))
# Final plot
comppops <- comppops %>% filter(pop2 != "Emilia-Romagna")
comppops <- comppops%>% rename(Population = "pop1")
comppops$Population <- as_factor(comppops$Population)
comppops$Population <- fct_recode(comppops$Population, Modena="M", "Emilia-Romagna (other)"="ERo", "Emilia-Romagna"="ER", Italy="I")
comppops$Population <- fct_recode(comppops$Population, Modena="M", "Emilia-Romagna (other)"="ERo", "Emilia-Romagna"="ER", Italy="I")
comppops$pop2 <- fct_recode(comppops$pop2, Modena="M", "Emilia-Romagna (other)"="ERo", "Emilia-Romagna"="ER", Italy="I")


rm(compkp, comppa, compab, ERovsER_BSI, MvER_all, MvI_bsi, MvER_bsi, ERovsI_BSI, ERovsER_all, ERovsER_BSI)



ggplot(comppops %>% filter(pop2 !="Emilia-Romagna"), aes(year, totdiff, group=Population, fill=Population,
                     color=Population))+
    geom_line()+
    geom_ribbon(aes(ymin = totlci, ymax = totuci), alpha = 0.2,
                color=NA)+
    geom_ribbon(aes(ymin = italylci, ymax = italyuci), alpha = 0.2,
                linetype = "dotted", fill=NA)+
    geom_line(aes(x= year, y=italytotdiff), linetype= "dashed")+
    facet_grid(org~pop2+site1, scales="free_y")+
  geom_hline(yintercept=0, linetype="dashed", color="darkgrey")+
  scale_y_continuous ("Difference in resistance, %R",
                      labels=scales::percent_format(accuracy=1))+
  scale_x_continuous ("Year", breaks=seq(2008,2018,2))+
  theme_pnaat()+
  scale_color_viridis_d(option="A", begin=0.3, end=0.8)+
  scale_fill_viridis_d(option="A", begin = 0.3, end=0.8)+
  theme(axis.text.x = element_text(angle = 45, hjust =1))+
  theme(legend.position = "top")+
  theme(legend.title=element_text(size=9))+
  coord_cartesian(ylim=c(-0.4,0.4))

ggsave("./Output/regnat_diffs.jpg", 
       height=15, width=20, units="cm")


```

## 2.6 Difference in infection site resistance trends
```{r message=FALSE, warning=FALSE, results='hide', cache= TRUE}
## Obtain differences using the smooth_diff function
mod_site_kp <- smooth_diff(kp_pop_GAMx, pdat, "BSI.M", "All clinical.M", "sitepop")

mod_site_pa <- smooth_diff(pa_pop_GAMx, pdat, "BSI.M", "All clinical.M", "sitepop")

mod_site_ab <- smooth_diff(ab_pop_GAMx, pdat, "BSI.M", "All clinical.M", "sitepop")

ero_site_kp <- smooth_diff(kp_pop_GAMx, pdat, "BSI.ERo", "All clinical.ERo", "sitepop")

ero_site_pa <- smooth_diff(pa_pop_GAMx, pdat, "BSI.ERo", "All clinical.ERo", "sitepop")  

ero_site_ab <- smooth_diff(ab_pop_GAMx, pdat, "BSI.ERo", "All clinical.ERo", "sitepop")


site_kp <- rbind(mod_site_kp, ero_site_kp) %>% mutate(org = "K.pneumoniae") 
site_pa <- rbind(mod_site_pa, ero_site_pa) %>% mutate(org = "P.aeruginosa")
site_ab <- rbind(mod_site_ab, ero_site_ab) %>% mutate(org = 
"A.baumannii")

sitediffs <- cbind(date = seq(2008, 2018, length = 264), 
             rbind(site_kp, site_pa, site_ab)) %>% 
  separate(col=pair, into=c("site1","pop1", "pop2"),
           sep = "[.]", remove = TRUE) %>% 
  separate(col="pop1", into=c("pop1", "null"), 
           sep = "[-]", remove = TRUE) %>% dplyr::select(-"null") %>% rename(Population = "pop1")
sitediffs$Population <- as_factor(sitediffs$Population)

sitediffs$Population <- fct_recode (sitediffs$Population, Modena = "M", "Emilia-Romagna (other)" = "ERo")


rm(site_kp, site_pa, site_ab)

#### plots
ggplot(sitediffs, aes(date, diff, group=Population, fill=Population,
                     color=Population))+
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2,
                color=NA)+
    geom_line()+
    facet_grid(org~Population, scales="free_y")+
  geom_hline(yintercept=0, linetype="dashed", color="darkgrey")+
  scale_y_continuous ("Relative difference (logit %R)")+
  scale_x_continuous ("Year", breaks=seq(2008,2018,2))+
  theme_pnaat()+
  scale_color_viridis_d(option="A", begin=0.8, end=0.2)+
  scale_fill_viridis_d(option="A", begin = 0.8, end=0.2)+
  theme(axis.text.x = element_text(angle = 45, hjust =1))+
  theme(legend.position = "none")+
  coord_cartesian (ylim=c(-2,2))

ggsave("./Output/regnat_diffs_site.jpg", 
       height=15, width=20, units="cm")

```
