---
title: "Modena Time Series Analysis"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  pdf_document:
    toc: yes
---

# Introduction
This R notebook provides code to reproduce results for the paper *Effects of antimicrobial stewardship and infection prevention and control on the incidence of carbapenem resistant gram-negative infections in Modena, Italy: a non-linear time-series analysis*. 
It requires installing of necessary packages


Data and code are available at the Github repository
[via this link](https://github.com/TimLawes/modena_carbapenem_resistance)

# 1. Preliminaries

## 1.1 Load required packages and customised plot theme
```{r, messages =FALSE, warning=FALSE}

library(tidyverse) # to enable tidyverse functions


library(earth) # for MARS
library(tseries) # for working with TS data
library(tsutils) #
library(forecast) # for forcasting
library(fpp3) # enables tsibble, and fable packages
library(plotmo) # plotting MARS model from earth
library(mgcv) # for GLM and GAM models
library(smooth)
library(lmtest) # comparing models 
library(stats)
library(grid)
library(lattice)
library(viridis) # for viridis colour palettes
library(hrbrthemes) #for plot themes
library(sysfonts) # for using google fonts
library(lubridate) # for working with datetime variables
library(car) # for Cook's distances
library(strucchange) # for detecting structural changes in time-series
library(ganttrify) # for time lines (Gantt type)
library(patchwork) # for combining different plot types
library(missForest) # for missing value imputation by random forest
library(mgcViz) # for plots and model checks of GAM
source("cochrane_plot_theme.R") # imports a custom ggplot theme defined in other script
source("OutlierTreatment2.R") # imports custom function for detecting extreme outliers and replacing.
```

## 1.3 Load data and process
```{r, messages =FALSE, warning=FALSE, results='hide'}

# Read in the main data 
main <- read_csv ("../Data/raw data/modena_final.csv", show_col_types =FALSE)

date <- yearmonth(seq(as.Date("2008-01-01"), as.Date("2018-12-31"), by = "1 month")) %>% as_tibble() %>% rename(date="value")

main <- cbind(date,main) %>% dplyr::select(!c(Month, study_m))

# Create wide as a time-series indexed table (tsibble)
tswide <- main %>% mutate(
  across(gmneg_all_bsi:crpa_cohca_swab,~./obd*10000),
  across(abhr: rectalswab, ~./obd*1000))

```

## 1.4 Outlier detection and time-series cleaning
```{r, messages=FALSE, warning=FALSE}


# outlier detection and replacement with cleaned values
wide_clean <- tswide %>% as_tibble %>% dplyr::select(date, obd, bedocc,abhr:all_monitored)

newabx.matrix<-rep(0,132)
for(i in 1:length(wide_clean[1,])){
abx<- wide_clean[,i]
newabx<-OutlierTreatment2(abx,freq=12,start.date=c(2008,1), only.clean=TRUE)
newabx.matrix <-cbind(newabx.matrix, newabx)
}

newabx.matrix<-newabx.matrix[,-1]

colnames(newabx.matrix)<-c(paste0(colnames(wide_clean)))
newabx.matrix<- newabx.matrix[,!colnames(newabx.matrix) %in% 'date']

tmp <-tswide %>% dplyr::select(!c(date, obd,bedocc,
abhr:all_monitored))

tswide_clean <- cbind(newabx.matrix,tmp)
rm(tmp, newab.matrix, date, abx, tswide, wide_clean)
```

