---
title: "Modena Time Series Analysis"
output:
  html_document:
    df_print: paged
    toc: yes
  word_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

# Introduction

This R notebook provides code to reproduce results for the paper *Effects of antimicrobial stewardship and infection prevention and control on the incidence of carbapenem resistant gram-negative infections in Modena, Italy: a non-linear time-series analysis*. It requires installing of necessary packages

Data and code are available at the Github repository [via this link](https://github.com/TimLawes/modena_carbapenem_resistance)

# 1. Preliminaries

## 1.1 Install copies of required packages to private library

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# Install renv for private package library if not available
if(!require(renv)){
    install.packages("renv",destdir = "C://R",dependencies = TRUE)
}

Sys.setenv(LOCALAPPDATA=renv::paths$library())
renv::restore()
renv::activate()
renv::snapshot()

```

## 1.2 Load required packages and customised plot theme

```{r warning=FALSE, include=FALSE, messages=FALSE}

# optional:  
# package.list<-c('tidyverse','earth','tseries','tsutils','forecast','fpp3','plotmo','mgcv','smooth','lmtest','stats','grid','lattice','viridis','hrbrthemes','sysfonts','lubridate','car','strucchange','ganttrify','patchwork','missForest','mgcViz','lmtest','feasts','fpp3','AID','devtools') 
# install.packages(package.list, destdir = "C://R",dependencies = TRUE)

library(tidyverse) # to enable tidyverse functions   install.packages("package",destdir = "C://R")
library(earth) # for MARS
library(tseries) # for working with TS data
library(tsutils) # install.packages("tsutils",destdir = "C://R")
library(forecast) # for forcasting
library(fpp3) # enables tsibble, and fable packages
library(plotmo) # plotting MARS model from earth
library(mgcv) # for GLM and GAM models
library(smooth) # install.packages("smooth",destdir = "C://R")
library(lmtest) # comparing models 
library(stats)
library(grid)
library(lattice) # install.packages("lattice",destdir = "C://R")
library(viridis) # for viridis colour palettes
library(hrbrthemes) #for plot themes
library(sysfonts) # for using google fonts
library(lubridate) # for working with datetime variables
library(car) # for Cook's distances
library(strucchange) # for detecting structural changes in time-series
library(ganttrify) # for time lines (Gantt type)
library(patchwork) # for combining different plot types
library(missForest) # for missing value imputation by random forest
library(mgcViz) # for plots and model checks of GAM
library(tsibble)
library(lmtest)
library(feasts)
library(fpp3)
library(AID)
library(readxl)
source("cochrane_plot_theme.R") # imports a custom ggplot theme defined in other script
source("OutlierTreatment2.R") # imports custom function for detecting extreme outliers and replacing.
```

## 1.3 Load data and process

```{r warning=FALSE, include=FALSE, messages=FALSE, results='hide'}

# Read in the main data 
main <- read_excel("../Data/raw data/modena_final_2008_2018.xlsx")

date <- yearmonth(seq(as.Date("2008-01-01"), as.Date("2018-12-31"), by = "1 month")) %>% as_tibble() %>% rename(date="value")

main <- cbind(date,main) %>% dplyr::select(!c(study_m, Month))


# Create wide as a time-series indexed table (tsibble)
tswide <- main %>% mutate(
  across(gmneg_all_bsi:crpa_cohca_swab,~./obds*10000),
  across(abhr: rectalswab, ~./obds*1000))  %>% 
  as_tsibble(index = date)

```

## 1.4 Outlier detection and time-series cleaning

```{r, messages=FALSE, warning=FALSE}


# outlier detection and replacement with cleaned values
wide_clean <- tswide %>% as_tibble %>% dplyr::select(date, obds, bedocc,abhr:all_monitor_abx)
                                                     
newabx.matrix<-rep(0,132)
for(i in 1:length(wide_clean[1,])){
abx<- wide_clean[,i]
newabx<-OutlierTreatment2(abx,freq=12,start.date=c(2008,1), only.clean=TRUE)
newabx.matrix <-cbind(newabx.matrix, newabx)
}

newabx.matrix<-newabx.matrix[,-1]

colnames(newabx.matrix)<-c(paste0(colnames(wide_clean)))
newabx.matrix<- newabx.matrix[,!colnames(newabx.matrix) %in% 'date']

tmp <-tswide %>% dplyr::select(!c(date, obds,bedocc,abhr:all_monitor_abx))

tswide_clean<-cbind(newabx.matrix,as.data.frame(tmp)) %>% as_tsibble(index=date)
#tswide_clean <- dplyr::bind_cols(newabx.matrix,tmp) %>% as_tsibble(index=date)
rm(tmp, newab.matrix, date, abx, tswide, wide_clean)

```

## 1.5 Trend decomposition

```{r, messages =FALSE, warning=FALSE}

# create long format data for incidence

tslong <- tswide_clean %>% dplyr::select(!c(hh_adherence:mrsa_all_bsi)) %>%
  pivot_longer(
  cols = csab_all_any:crpa_cohca_swab,
  names_to = c("IR", "org", "pop", "site"),
  names_pattern = "(cr|cs)(ab|kp|pa)_(.*)_(.*)",
  values_to = "incidence") %>% filter(pop!="NA") 

tslong<- tslong %>% mutate (bugs = paste(IR,org,pop,site))

tslong <- tslong %>% dplyr::select(-bedocc) %>% group_by(bugs,IR,org,pop,site) %>% nest() 

tslongbugs<- tslong %>% filter(pop=="hca" | pop=="conothca")

decompose <- function(data, y, ...){
    y <- enquo(y)
    dvar <- data %>% pull(!!y) %>% as.ts()
    tmpd <- data %>% model(STL(dvar ~ trend(window = 7) + 
                                 season(window =13), robust = TRUE)) %>%
      components()
    data <-cbind(data,tmpd[,3:6])
}

ts_decomp <- tslongbugs %>% 
    mutate(decs = map(data, decompose, y = incidence)) %>%
    dplyr::select(-data) %>% unnest(decs) %>% ungroup()
ts_decomp <- ts_decomp %>% mutate (trend=if_else(trend <0, 0, trend))

```

# 3. ARIMAs

In addition to time-series decomposition applied to all variables (to facilitate descriptive statistics) we further characterised key time-series, including:

-   Primary outcomes
-   Antibiotic time-series - for counterfactuals
-   Other infection time-series driving antibiotic use

## 3.1 Primary outcomes

The primary study outcomes of interest are:

-   crpa hca clinical,
-   crpa conothca clinical,
-   crkp hca clinical,
-   crkp conothca clinical,
-   crab hca clinical,
-   crab conothca clinical

### 3.1.1 Redefine timeseries datasets for ARIMAs

```{r message=FALSE, warning=FALSE, include=FALSE}

# Derive pre-intervention time-series
ts_p1_pre_ihh_itargscr <- tswide_clean %>% filter (phase <=1)

ts_p2_pre_icarba <- tswide_clean %>% filter (phase <=2)
date_icarba<-ts_p2_pre_icarba$date

ts_p3_pre_iabs_iunivscr <- tswide_clean %>% filter (phase <=3)

##### auto arima function for reference

auto.arima.df <- function(data, y, ...){
    y <- enquo(y)
    yts <- data %>% 
        pull(!!y) %>% 
        as.ts()
    auto.arima(yts, ...)
}

```

### 3.1.2 HCA-CRKp clinical infections (*crkp hca clinical*)

```{r message=FALSE, warning=FALSE, include=FALSE}

#### Plot data, ACF, PACF

tswide_clean <- tswide_clean %>% as_tsibble(index=date)

tswide_clean %>% gg_tsdisplay(crkp_ho_clinical, "partial")


#### Apply a unit root test to check if time-series stationary,  determine number of non-seasonal (ndiffs) and seasonal (nsdiffs) differences required to make stationary, and plot differenced time-series if required.
tswide_clean %>% features(crkp_ho_clinical, unitroot_kpss)
tswide_clean %>% features(crkp_ho_clinical, unitroot_ndiffs) 
tswide_clean %>% features(crkp_ho_clinical, unitroot_nsdiffs,.period=12) 

#### Examine need for a Box-Cox data transformation to reduce heteroskedasticity (i.e variability in the variance of the time-series)
lambda<-guerrero(tswide_clean$crkp_ho_clinical, .period=12L)

#### create new transformed variable
tswide_clean$crkp_ho_clinical_bc <- (((tswide_clean$crkp_ho_clinical)^lambda)-1)/lambda

####  Plot the transformed variable, ACF, PACF
tswide_clean %>% gg_tsdisplay(crkp_ho_clinical_bc, "partial")

#### Repeat unit root tests and differencing number evaluations on the transformed data
tswide_clean %>% features(crkp_ho_clinical_bc, unitroot_kpss)
tswide_clean %>% features(crkp_ho_clinical_bc, unitroot_ndiffs)
tswide_clean %>% features(crkp_ho_clinical_bc, unitroot_nsdiffs,.period=12) 

tswide_clean %>% gg_tsdisplay(difference(crkp_ho_clinical_bc), "partial")  

# fit best manual and auto ARIMA with default approach for NON-TRANSFORMED DATA
arima_crkp_ho <- tswide_clean %>% model(
  arima_crkp_ho_101 = ARIMA(crkp_ho_clinical ~ pdq(1,0,1)),
  stepwise_crkp_ho = ARIMA(crkp_ho_clinical)) 

glance(arima_crkp_ho) %>% arrange(AICc) %>% dplyr::select(.model:BIC)
arima_crkp_ho  %>% dplyr::select(arima_crkp_ho_101) %>% gg_tsresiduals() # compare models
arima_crkp_ho %>% dplyr::select(stepwise_crkp_ho) %>% gg_tsresiduals() # compare models


#### Fit best manual and auto ARIMA on transformed data
arima_crkp_ho_bc <- tswide_clean %>% model(
  arima_crkp_ho_bc_111 = ARIMA(crkp_ho_clinical_bc ~ pdq(1,1,1)),
  arima_crkp_ho_bc_101_100 = ARIMA(crkp_ho_clinical_bc ~ pdq(1,0,1) + PDQ(1,0,0)),
    arima_crkp_ho_bc_011_100 = ARIMA(crkp_ho_clinical_bc ~ pdq(0,1,1) + PDQ(1,0,0)),
  arima_crkp_ho_bc_101 = ARIMA(crkp_ho_clinical_bc ~ pdq(1,0,1)),
  arima_crkp_ho_bc_001 = ARIMA(crkp_ho_clinical_bc ~ pdq(0,0,1)),
  stepwise_crkp_ho_bc = ARIMA(crkp_ho_clinical_bc)) 

glance(arima_crkp_ho_bc) %>% arrange(AICc) %>% dplyr::select(.model:BIC)

arima_crkp_ho_bc  %>% dplyr::select(arima_crkp_ho_bc_011_100) %>% gg_tsresiduals() # compare models
arima_crkp_ho_bc %>% dplyr::select(stepwise_crkp_ho_bc) %>% gg_tsresiduals() # compare models

#### Check portmanteau tests of best fit option
augment(arima_crkp_ho_bc) %>% 
  filter(.model=='arima_crkp_ho_bc_011_100') %>%
  features(.innov, ljung_box, lag = 10, dof = 3)  # portmanteau test on best manual



```

### 3.1.3 HCA-CRAb clinical infections (*crab hca clinical*)

```{r message=FALSE, warning=FALSE, include=FALSE}

#### Plot data, ACF, PACF
tswide_clean %>% gg_tsdisplay(crab_ho_clinical, "partial") 

#### Apply a unit root test to check if time-series stationary,  determine number of non-seasonal (ndiffs) and seasonal (nsdiffs) differences required to make stationary, and plot differenced time-series if required.
tswide_clean %>% features(crab_ho_clinical, unitroot_kpss)
tswide_clean %>% features(crab_ho_clinical, unitroot_ndiffs)
tswide_clean %>% features(crab_ho_clinical, unitroot_nsdiffs,.period=12)
tswide_clean %>% gg_tsdisplay(difference(crab_ho_clinical), "partial") 


#### Examine need for a Box-Cox data transformation to reduce heteroskedasticity (i.e variability in the variance of the time-series)
lambda<-guerrero(tswide_clean$crab_ho_clinical, .period=12L)

#### Fit best manual and auto ARIMA on transformed data
arima_crab_ho <- tswide_clean %>% model(
  arima_crab_ho_111 = ARIMA(crab_ho_clinical ~ pdq(1,1,1)),
  arima_crab_ho_011 = ARIMA(crab_ho_clinical ~ pdq(0,1,1)),
  arima_crab_ho_110 = ARIMA(crab_ho_clinical ~ pdq(1,1,0)),
  arima_crab_ho_101 = ARIMA(crab_ho_clinical ~ pdq(1,0,1)),
  stepwise_crab_ho = ARIMA(crab_ho_clinical))

glance(arima_crab_ho) %>% arrange(AICc) %>% dplyr::select(.model:BIC)

# check the residuals of best models to ensure white noise.
arima_crab_ho  %>% dplyr::select(arima_crab_ho_011) %>% gg_tsresiduals() 
arima_crab_ho %>% dplyr::select(arima_crab_ho_101) %>% gg_tsresiduals()

#### Check portmanteau tests of best fit option
augment(arima_crab_ho) %>% 
  filter(.model=='arima_crab_ho_011') %>%
  features(.innov, ljung_box, lag = 10, dof = 3)  # portmanteau test on best manual


```

### 3.1.4 HCA-CRPa clinical infections (*crpa hca clinical*)

```{r message=FALSE, warning=FALSE, include=FALSE}

#### Plot data, ACF, PACF
tswide_clean %>% gg_tsdisplay(crpa_ho_clinical, "partial")

#### Apply a unit root test to check if time-series stationary,  determine number of non-seasonal (ndiffs) and seasonal (nsdiffs) differences required to make stationary, and plot differenced time-series if required.
tswide_clean %>% features(crpa_ho_clinical, unitroot_kpss)
tswide_clean %>% features(crpa_ho_clinical, unitroot_ndiffs) 
tswide_clean %>% features(crpa_ho_clinical, unitroot_nsdiffs,.period=12) 

tswide_clean %>% gg_tsdisplay(difference(crpa_ho_clinical), "partial") 
crpa_ho_clin_diff1 <-  tsibble::difference(tswide_clean$crpa_ho_clinical, lag =1, differences = 1)
tswide_clean %>% gg_tsdisplay(crpa_ho_clin_diff1, "partial")

#### Examine need for a Box-Cox data transformation to reduce heteroskedasticity (i.e variability in the variance of the time-series)
tswide_clean <- tswide_clean %>% dplyr::mutate(crpa_ho_clin_diff1 = replace_na(crpa_ho_clin_diff1, 0))
tswide_clean <- tswide_clean %>% dplyr::mutate(new_crpa_ho_clin_diff1 = crpa_ho_clin_diff1+10)
boxcoxnc(tswide_clean$new_crpa_ho_clin_diff1, method = "sw", lambda2 = 10)


#### Repeat unit root tests and differencing number evaluations on the transformed data
tswide_clean %>% features(crpa_ho_clin_diff1, unitroot_kpss)
tswide_clean %>% features(crpa_ho_clin_diff1, unitroot_ndiffs)
tswide_clean %>% features(crpa_ho_clin_diff1, unitroot_nsdiffs, .period=12)

#### Fit best manual and auto ARIMA
arima_crpa_ho <- tswide_clean %>% model(
  arima_crpa_ho_111 = ARIMA(crpa_ho_clinical ~ pdq(1,1,1)),
  arima_crpa_ho_011 = ARIMA(crpa_ho_clinical ~ pdq(0,1,1)),
  arima_crpa_ho_101 = ARIMA(crpa_ho_clinical ~ pdq(1,0,1)),
  stepwise_crpa_ho = ARIMA(crpa_ho_clinical)) 

glance(arima_crpa_ho) %>% arrange(AICc) %>% dplyr::select(.model:BIC)

#### Examine residuals of ARIMA to check correct identification
arima_crpa_ho  %>% dplyr::select(arima_crpa_ho_011) %>% gg_tsresiduals()
arima_crpa_ho %>% dplyr::select(arima_crpa_ho_101) %>% gg_tsresiduals()
arima_crpa_ho %>% dplyr::select(stepwise_crpa_ho) %>% gg_tsresiduals() 

#### Check portmanteau tests of best fit option
augment(arima_crpa_ho) %>% 
  filter(.model=='arima_crpa_ho_101') %>%
  features(.innov, ljung_box, lag = 10, dof = 3)  # portmanteau test on best manual
```

## 3.2 Antibiotic and IPC time series

```{r}

# Nb: we use the pre-intervention dataset "ts_p2_pre_icarba" for fitting the baseline ARIMA with exogenous variables (ARIMAX) for carbapenems. For other antibiotics we should use the  "ts_p3_pre_iabs_iunivscr" data
pre_icarba <- ts_p2_pre_icarba %>% as_tsibble(index="date")
pre_iabs <- ts_p3_pre_iabs_iunivscr %>% as_tsibble(index="date")



```

## 3.3 Influential infection time-series

These time-series are expected to be related to antibiotic use in the population. Bidirectional relationships may exist but we hypothesised that hospital-wide use of important antibiotic exposures driving carbapenem resistance may be responsive to the following:

-   All gram-negative bloodstream isolates (BSIs) (*gmneg_all_bsi*)
-   Carbapenem and cephalosporin sensitive E.coli BSI (*ecoliccs_all_bsi*)
-   3rd generation cephalosporins resistant E.coli BSI (*ecoli3gcr_all_bsi*)

### 3.3.1 Gram-Neg BSIs

```{r}
### Carbapenems######################

# Prepare differenced time-series
pre_icarba$gmneg_diff <-  tsibble::difference(pre_icarba$gmneg_all_bsi, lag =1, differences = 1)
pre_icarba$carbapenem_diff <-  tsibble::difference(pre_icarba$carbapenem, lag =1, differences = 1)
pre_icarba_ed2 <- pre_icarba [2:72,] %>% as_tsibble(index="date")
pre_icarba_ed4 <- pre_icarba [4:72,] %>% as_tsibble(index="date")

# Explore cross-correlations with raw data
CCF(pre_icarba, carbapenem, gmneg_all_bsi, lag_max=24) %>% autoplot(level=95)

# Explore cross-correlations with differenced data
CCF(pre_icarba_ed2, carbapenem_diff, gmneg_diff, lag_max=24)%>% autoplot(level=95)

# Identify an ARIMAX model for carbapenems 
carba_pre1 <- pre_icarba %>% model(ARIMA(carbapenem~ -1 + pdq(1,1,0)+PDQ(0,0,0)+lag(gmneg_all_bsi,2)))
report(carba_pre1)
carba_pre1 %>%  gg_tsresiduals() 

carba_pre2 <- pre_icarba_ed2 %>% model(ARIMA(carbapenem  ~ -1 + pdq(1,1,0)+PDQ(0,0,0)+lag(gmneg_diff,2)))
report(carba_pre2)
carba_pre2 %>%  gg_tsresiduals() 

# Looking at the correlation between gm neg BSI and differenced gm neg BSI (y-axis)
gmneg_bsi<- pre_icarba$gmneg_all_bsi[1:71]
differenced_gm_negbsi <- pre_icarba$gmneg_diff [2:72]
gmneg <- cbind(gmneg_bsi,differenced_gm_negbsi )
gmneg <-gmneg %>% as_tibble()
cor(gmneg$gmneg_bsi, gmneg$differenced_gm_negbsi)
ggplot(gmneg, aes(gmneg_bsi,differenced_gm_negbsi))+geom_point()

```

```{r}
### Fluorquinolones ######################

# Prepare differenced time-series
# pre_icarba$gmneg_diff <-  tsibble::difference(pre_icarba$gmneg_all_bsi, lag =1, differences = 1)
pre_icarba$fq_diff <-  tsibble::difference(pre_icarba$fq, lag =1, differences = 1)
pre_icarba_ed2 <- pre_icarba [2:72,] %>% as_tsibble(index="date")
pre_icarba_ed4 <- pre_icarba [4:72,] %>% as_tsibble(index="date")

# Explore cross-correlations with raw data
CCF(pre_icarba, fq, gmneg_all_bsi, lag_max=24) %>% autoplot(level=95)

# Explore cross-correlations with differenced data
CCF(pre_icarba_ed2, fq_diff, gmneg_diff, lag_max=24)%>% autoplot(level=95)

# Identify an ARIMAX model for carbapenems 

fq_pre1 <- pre_icarba %>% model(ARIMA(fq~ -1 + pdq(0,1,1)+PDQ(0,0,0)+lag(gmneg_all_bsi,1) ))
report(fq_pre1)
fq_pre1 %>%  gg_tsresiduals() 

fq_pre2 <- pre_icarba_ed2 %>% model(ARIMA(fq  ~ -1 + pdq(0,1,1)+PDQ(0,0,0)+lag(gmneg_diff,1)))
report(fq_pre2)
fq_pre2 %>%  gg_tsresiduals() 


```


### 3.3.2 Forecasting ABX with Gram-Neg BSIs

```{r}

adf.test(pre_icarba$carbapenem)
unitroot_ndiffs(pre_icarba$carbapenem)
auto.arima(pre_icarba$carbapenem,allowdrift = TRUE, allowmean=TRUE,max.q = 0,xreg = cbind(lag(pre_icarba$gmneg_all_bsi,2)))

model.pre_icarba.carba<-Arima(pre_icarba$carbapenem, order = c(1,0,0), include.drift = TRUE,xreg = lag(pre_icarba$gmneg_all_bsi,2))
coeftest(model.pre_icarba.carba)
Acf(model.pre_icarba.carba$residuals)
AIC(model.pre_icarba.carba)

gmneg_all_future  <-tswide_clean$gmneg_all_bsi[73:132]

fore.icarba.carba<-forecast(model.pre_icarba.carba,h = 60, xreg = gmneg_all_future) 
autoplot(fore.icarba.carba)


```
```{r}

adf.test(pre_icarba$fq)
unitroot_ndiffs(pre_icarba$fq)
auto.arima(pre_icarba$fq,allowdrift = TRUE,max.q = 0,xreg = cbind(lag(pre_icarba$gmneg_all_bsi,1)))

model.pre_icarba.fq<-Arima(pre_icarba$fq,order = c(1,0,0),include.drift = TRUE, xreg = cbind(lag(pre_icarba$gmneg_all_bsi,1)))
summary(model.pre_icarba.fq)
coeftest(model.pre_icarba.fq)
Acf(model.pre_icarba.fq$residuals)
AIC(model.pre_icarba.fq)

gmneg_all_future  <-tswide_clean$gmneg_all_bsi[73:132]

fore.icarba.fq<-forecast(model.pre_icarba.fq,h = 60, xreg = gmneg_all_future) 
autoplot(fore.icarba.fq)


```

```{r}

adf.test(pre_icarba$x3gc)
unitroot_ndiffs(pre_icarba$x3gc)
auto.arima(pre_icarba$x3gc)
auto.arima(pre_icarba$x3gc,allowdrift = TRUE,max.q = 0,xreg = cbind(lag(pre_icarba$gmneg_all_bsi,4)))
model.pre_icarba.x3gc<-Arima(pre_icarba$x3gc,order = c(1,0,0), include.drift = TRUE, xreg = lag(pre_icarba$gmneg_all_bsi,4))
summary(model.pre_icarba.x3gc)
coeftest(model.pre_icarba.x3gc)
Acf(model.pre_icarba.x3gc$residuals)
AIC(model.pre_icarba.x3gc)

gmneg_all_future  <-tswide_clean$gmneg_all_bsi[73:132]

fore.icarba.x3gc<-forecast(model.pre_icarba.x3gc,h = 60, xreg = gmneg_all_future) 
autoplot(fore.icarba.x3gc)


```
# 4. Interactions and VARs

First we estimate VAR gram negatve bsi, ecoliccs and ecoli3gcr with each ABX

```{r}
### VAR evaluation function
library(lmtest)  # install.packages("lmtest")
library(vars)    # install.packages("vars")


# Defining function for VAR evaluation 2 variables y and series
eval_var<-function(y, series, names){

  # Make series to be stationary
  
dy=0; dseries=0

if(unitroot_ndiffs(y)!=0){dy<-unitroot_ndiffs(y);y<-diff(y, differences=dy)}
if(unitroot_ndiffs(series)!=0){dseries<-unitroot_ndiffs(series); series<-diff(series, differences=dseries)}

if(dy>dseries){series<-head(series,-I(dy-dseries))} 
if(dy<dseries){y<-head(y,-I(dseries-dy))}    

assign(names[1],y)
assign(names[2],series)

  # Select order and estimate VAR
matrix<-cbind(y,series)
colnames(matrix)<-names

var.fit <- vars::VAR(matrix,lag.max = 4)

  # Some test form var estimation
serial.test<-serial.test(var.fit, lags.pt=10, type="PT.asymptotic")

normality.test <- normality.test(var.fit)

var.stability <- stability(var.fit, type = "Rec-CUSUM")

   # Impulse response function

impulse.response <- vars::irf(var.fit , n.ahead = 6, ortho = TRUE,
               cumulative = F, boot = TRUE, ci = 0.95, runs = 100)

  # Variance decomposition of the forecasts

var.decomp<-fevd(var.fit)

  # causality analysis

causality<- list(cause.y=causality(var.fit,  cause=names[1],boot=TRUE, boot.runs=1000)$Granger, cause.series=causality(var.fit, cause=names[2], boot=TRUE, boot.runs=1000)$Granger)


return(
  list(
    var.fit=var.fit,
    causality=causality,
    impulse.response=impulse.response,
    serial.test=serial.test,
    normality.test=normality.test, 
    var.stability=var.stability,
    var.decomp=var.decomp
    )
  )
}



### VAR Gram Negative, ecoliccs ecoli3gcr( Pre_icarba[,18:20]) => abx( Pre_icarba[,5:9]) 


for (n in 18:20){
  
 for(j in 5:9){
names<-colnames(pre_icarba)[c(n,j)]
var<-eval_var(y=pre_icarba[,n], 
              series=pre_icarba[,j],
              names=names)

#VAR.data<-data.frame(pre_icarba[,n], pre_icarba[,j])
#var<-VAR(VAR.data, ic="SC",lag.max = 12)
#irf<-irf(var,n.ahead = 12,ortho=TRUE,cumulative=FALSE,boot=TRUE, ci=c(0.95), runs=50)
#plot(irf)

plot(var$impulse.response)

var$causality$cause.y
var$causality$cause.series

# optional: plot(var$var.decomp)

assign(paste0(c("var",colnames(pre_icarba)[n],colnames(pre_icarba)[j]), sep=".", collapse = ""),
       var)
 }
}


```

Estimate VAR between swab and clinical
```{r}

pathogen<-c("crab","crkp","crpa")
pop<-c("ho","cohca","conothca", "hca")


for (n in 1:3){
  
 for(j in 1:4){

names.var<-c(paste("pre_icarba$", paste(pathogen[n],pop[j],"swab",sep="_"), sep=""),paste("pre_icarba$", paste(pathogen[n],pop[j],"clinical",sep="_"), sep=""))

names<-c(paste(pathogen[n],pop[j],"swab",sep="_"),paste(pathogen[n],pop[j],"clinical",sep="_"))
names0<-c("swab","clinical")   
   

var<-eval_var(y=eval( parse(text=names.var[1])), 
              series=eval( parse(text=names.var[2])),
              names=names0)
cat(paste(names[1],names[2],sep="<=>"))
plot(var$impulse.response) # It is necessary to put irf in one plot and common title for <org> <pop>

var$causality$cause.y
var$causality$cause.series

# optional: plot(var$var.decomp)

assign(paste0(c("var",names[2],names[1]), sep=".", collapse = ""),var)
 }
}

```

Interspecies competition

```{r}

eval_var3<-function(x,y,z,names){


  # Make series to be stationary
  
dx=0; dy=0 ; dz=0
if(unitroot_ndiffs(x)!=0){dx<-unitroot_ndiffs(x);x<-diff(x, differences=dx)}
if(unitroot_ndiffs(y)!=0){dy<-unitroot_ndiffs(y);y<-diff(y, differences=dy)}
if(unitroot_ndiffs(z)!=0){dz<-unitroot_ndiffs(z);z<-diff(z, differences=dz)}

maxd<-max(dx,dy,dz)

if(dx<maxd){x<-head(x,-I(maxd-dx))} 
if(dy<maxd){y<-head(y,-I(maxd-dy))}    
if(dz<maxd){z<-head(z,-I(maxd-dz))}    


assign(names[1],x)
assign(names[2],y)
assign(names[3],z)

  # Select order and estimate VAR
matrix<-cbind(x,y,z)
colnames(matrix)<-names

var.fit <- VAR(matrix, lag.max =4)

  # Some test form var estimation

serial.test<-serial.test(var.fit, lags.pt=10, type="PT.asymptotic")

normality.test <- normality.test(var.fit)

var.stability <- stability(var.fit, type = "Rec-CUSUM")
   
  # Impulse response function

impulse.response <- irf(var.fit , n.ahead = 6, ortho = TRUE,
               cumulative = F, boot = TRUE, ci = 0.95, runs = 100)

  # Variance decomposition of the forescats

var.decomp<-fevd(var.fit)

# causality analisys

causality<- list(
  cause.x=causality(var.fit,  cause=names[1],boot=TRUE, boot.runs=1000)$Granger, 
  cause.y=causality(var.fit,  cause=names[2],boot=TRUE, boot.runs=1000)$Granger,
  cause.z=causality(var.fit,  cause=names[3],boot=TRUE, boot.runs=1000)$Granger)
  

return(
  list(
    var.fit=var.fit,
    causality=causality,
    impulse.response=impulse.response,
    serial.test=serial.test,
    normality.test=normality.test, 
    var.stability=var.stability,
    var.decomp=var.decomp
    )
  )
}

names<-pathogen

for( j in 1:4){

names.var<-vector()
for (n in 1:3){
names.var<-c(names.var, paste("pre_icarba$", paste(pathogen[n],pop[j],"clinical",sep="_"), sep=""))
}

x=eval( parse(text=names.var[1])) 
y=eval( parse(text=names.var[2]))
z=eval( parse(text=names.var[3]))

var<-eval_var3(x=x,y=y,z=z, names=names)

cat(paste(names[1],names[2],names[3],sep="<=>"))
plot(var$impulse.response)


var$causality$cause.x
var$causality$cause.y
var$causality$cause.z

plot(var$serial.test)
plot(var$normality.test)
plot(var$var.stability)
# optional: plot(var$var.decomp)

assign(paste0(c("var",names[1],names[2], names[3], pop[j]), sep=".", collapse = ""),var)

}

```

Interaction among sensitive and resistant strains

```{r}

pathogen0<-rbind(c("crab","crkp","crpa"),c("csab","cskp","cspa"))

pop<-c("ho","cohca","conothca", "hca")


for (n in 1:3){
  
 for(j in 1:4){

names.var<-c(paste("pre_icarba$", paste(pathogen0[1,n],pop[j],"clinical",sep="_"), sep=""),paste("pre_icarba$", paste(pathogen0[2,n],pop[j],"clinical",sep="_"), sep=""))

names<-c(paste(pathogen0[1,n],pop[j],"clinical",sep="_"),paste(pathogen0[2,n],pop[j],"clinical",sep="_"))
 
var<-eval_var(y=eval( parse(text=names.var[1])), 
              series=eval( parse(text=names.var[2])),
              names=names)

cat(paste(names[1],names[2],sep="<=>"))
plot(var$impulse.response) # It is necessary to put irf in one plot and common title for <org> <pop>

var$causality$cause.y
var$causality$cause.series

plot(var$var.decomp)

assign(paste0(c("var",names[2],names[1]), sep=".", collapse = ""),var)
 }
}

```








```{r}
library(devtools)
install_github("MatthieuStigler/tsDyn", ref="Dev94", subdir="tsDyn")  
library(tsDyn)
```

```{r}

```

# 5. GAMs

```{r}

```

# 6. MARS

```{r}

```
