---
title: "Modena Time Series Analysis"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  pdf_document:
    toc: yes
---

# Introduction
This R notebook provides code to reproduce results for the paper *Effects of antimicrobial stewardship and infection prevention and control on the incidence of carbapenem resistant gram-negative infections in Modena, Italy: a non-linear time-series analysis*. 
It requires installing of necessary packages


Data and code are available at the Github repository
[via this link](https://github.com/TimLawes/modena_carbapenem_resistance)

# 1. Preliminaries
## 1.1 Install copies of required packages to private library
```{r message=FALSE, warning=FALSE, include=FALSE}

# Install renv for private package library if not available
if(!require(renv)){
    install.packages("renv")
}

Sys.setenv(LOCALAPPDATA=renv::paths$library())
renv::restore()
renv::activate()

```

## 1.2 Load required packages and customised plot theme
```{r, messages =FALSE, warning=FALSE}

library(tidyverse) # to enable tidyverse functions
library(earth) # for MARS
library(tseries) # for working with TS data
library(tsutils) #
library(forecast) # for forcasting
library(fpp3) # enables tsibble, and fable packages
library(plotmo) # plotting MARS model from earth
library(mgcv) # for GLM and GAM models
library(smooth)
library(lmtest) # comparing models 
library(stats)
library(grid)
library(lattice)
library(viridis) # for viridis colour palettes
library(hrbrthemes) #for plot themes
library(sysfonts) # for using google fonts
library(lubridate) # for working with datetime variables
library(car) # for Cook's distances
library(strucchange) # for detecting structural changes in time-series
library(ganttrify) # for time lines (Gantt type)
library(patchwork) # for combining different plot types
library(missForest) # for missing value imputation by random forest
library(mgcViz) # for plots and model checks of GAM
source("cochrane_plot_theme.R") # imports a custom ggplot theme defined in other script
source("OutlierTreatment2.R") # imports custom function for detecting extreme outliers and replacing.
```

## 1.3 Load data and process
```{r, messages =FALSE, warning=FALSE, results='hide'}

# Read in the main data 
main <- read_csv ("../Data/raw data/modena_final.csv", show_col_types =FALSE)

date <- yearmonth(seq(as.Date("2008-01-01"), as.Date("2018-12-31"), by = "1 month")) %>% as_tibble() %>% rename(date="value")

main <- cbind(date,main) %>% dplyr::select(!c(Month, study_m))

# Create wide as a time-series indexed table (tsibble)
tswide <- main %>% mutate(
  across(gmneg_all_bsi:crpa_cohca_swab,~./obd*10000),
  across(abhr: rectalswab, ~./obd*1000))  %>% 
  as_tsibble(index = date)

```

## 1.4 Outlier detection and time-series cleaning
```{r, messages=FALSE, warning=FALSE}


# outlier detection and replacement with cleaned values
wide_clean <- tswide %>% as_tibble %>% dplyr::select(date, obd, bedocc,abhr:all_monitored)

newabx.matrix<-rep(0,132)
for(i in 1:length(wide_clean[1,])){
abx<- wide_clean[,i]
newabx<-OutlierTreatment2(abx,freq=12,start.date=c(2008,1), only.clean=TRUE)
newabx.matrix <-cbind(newabx.matrix, newabx)
}

newabx.matrix<-newabx.matrix[,-1]

colnames(newabx.matrix)<-c(paste0(colnames(wide_clean)))
newabx.matrix<- newabx.matrix[,!colnames(newabx.matrix) %in% 'date']

tmp <-tswide %>% dplyr::select(!c(date, obd,bedocc,
abhr:all_monitored))

tswide_clean <- cbind(newabx.matrix,tmp) %>% as_tsibble(index=date)
rm(tmp, newab.matrix, date, abx, tswide, wide_clean)
```


## 1.5 Trend decomposition
```{r, messages =FALSE, warning=FALSE}

# create long format data for incidence

tslong <- tswide_clean %>% dplyr::select(!c(hh_adherence:mrsa_all_bsi)) %>%
  pivot_longer(
  cols = csab_all_any:crpa_cohca_swab,
  names_to = c("IR", "org", "pop", "site"),
  names_pattern = "(cr|cs)(ab|kp|pa)_(.*)_(.*)",
  values_to = "incidence") %>% filter(pop!="NA") 

tslong<- tslong %>% mutate (bugs = paste(IR,org,pop,site))

tslong <- tslong %>% dplyr::select(-bedocc) %>% group_by(bugs,IR,org,pop,site) %>% nest() 

tslongbugs<- tslong %>% filter(pop=="hca" | pop=="conothca")

decompose <- function(data, y, ...){
    y <- enquo(y)
    dvar <- data %>% pull(!!y) %>% as.ts()
    tmpd <- data %>% model(STL(dvar ~ trend(window = 7) + 
                                 season(window =13), robust = TRUE)) %>%
      components()
    data <-cbind(data,tmpd[,3:6])
}

ts_decomp <- tslongbugs %>% 
    mutate(decs = map(data, decompose, y = incidence)) %>%
    dplyr::select(-data) %>% unnest(decs) %>% ungroup()
ts_decomp <- ts_decomp %>% mutate (trend=if_else(trend <0, 0, trend))

```


# 3. ARIMAs
In addition to time-series decomposition applied to all variables (to facilitate descriptive statistics) we further characterised key time-series, including:

* Primary outcomes
* Antibiotic time-series - for counterfactuals
* Other infection time-series driving antibiotic use


## 3.1 Primary outcomes
The primary study outcomes of interest are:

* crpa hca clinical, 
* crpa conothca clinical, 
* crkp hca clinical, 
* crkp conothca clinical, 
* crab hca clinical, 
* crab conothca clinical

### 3.1.1 Redefine timeseries datasets for ARIMAs
```{r message=FALSE, warning=FALSE, include=FALSE}

# Derive pre-intervention time-series
ts_p1_pre_ihh_itargscr <- tswide_clean %>% filter (phase <=1)
ts_p2_pre_icarba <- tswide_clean %>% filter (phase <=2)
ts_p3_pre_iabs_iunivscr <- tswide_clean %>% filter (phase <=3)

##### auto arima function for reference

auto.arima.df <- function(data, y, ...){
    y <- enquo(y)
    yts <- data %>% 
        pull(!!y) %>% 
        as.ts()
    auto.arima(yts, ...)
}

```

### 3.1.2 HCA-CRKp clinical infections (*crkp hca clinical*)
```{r message=FALSE, warning=FALSE, include=FALSE}

#### Plot data, ACF, PACF
tswide_clean %>% gg_tsdisplay(crkp_ho_clinical, "partial") 

#### Apply a unit root test to check if time-series stationary,  determine number of non-seasonal (ndiffs) and seasonal (nsdiffs) differences required to make stationary, and plot differenced time-series if required.
tswide_clean %>% features(crkp_ho_clinical, unitroot_kpss)
tswide_clean %>% features(crkp_ho_clinical, unitroot_ndiffs) 
tswide_clean %>% features(crkp_ho_clinical, unitroot_nsdiffs,.period=12) 

#### Examine need for a Box-Cox data transformation to reduce heteroskedasticity (i.e variability in the variance of the time-series)
lambda<-guerrero(tswide_clean$crkp_ho_clinical, .period=12L)

#### create new transformed variable
tswide_clean$crkp_ho_clinical_bc <- (((tswide_clean$crkp_ho_clinical)^lambda)-1)/lambda

####  Plot the transformed variable, ACF, PACF
tswide_clean %>% gg_tsdisplay(crkp_ho_clinical_bc, "partial")

#### Repeat unit root tests and differencing number evaluations on the transformed data
tswide_clean %>% features(crkp_ho_clinical_bc, unitroot_kpss)
tswide_clean %>% features(crkp_ho_clinical_bc, unitroot_ndiffs)
tswide_clean %>% features(crkp_ho_clinical_bc, unitroot_nsdiffs,.period=12) 

tswide_clean %>% gg_tsdisplay(difference(crkp_ho_clinical_bc), "partial")  

# fit best manual and auto ARIMA with default approach for NON-TRANSFORMED DATA
arima_crkp_ho <- tswide_clean %>% model(
  arima_crkp_ho_101 = ARIMA(crkp_ho_clinical ~ pdq(1,0,1)),
  stepwise_crkp_ho = ARIMA(crkp_ho_clinical)) 

glance(arima_crkp_ho) %>% arrange(AICc) %>% select(.model:BIC)
arima_crkp_ho  %>% select(arima_crkp_ho_101) %>% gg_tsresiduals() # compare models
arima_crkp_ho %>% select(stepwise_crkp_ho) %>% gg_tsresiduals() # compare models


#### Fit best manual and auto ARIMA on transformed data
arima_crkp_ho_bc <- tswide_clean %>% model(
  arima_crkp_ho_bc_111 = ARIMA(crkp_ho_clinical_bc ~ pdq(1,1,1)),
  arima_crkp_ho_bc_101_100 = ARIMA(crkp_ho_clinical_bc ~ pdq(1,0,1) + PDQ(1,0,0)),
    arima_crkp_ho_bc_011_100 = ARIMA(crkp_ho_clinical_bc ~ pdq(0,1,1) + PDQ(1,0,0)),
  arima_crkp_ho_bc_101 = ARIMA(crkp_ho_clinical_bc ~ pdq(1,0,1)),
  arima_crkp_ho_bc_001 = ARIMA(crkp_ho_clinical_bc ~ pdq(0,0,1)),
  stepwise_crkp_ho_bc = ARIMA(crkp_ho_clinical_bc)) 

glance(arima_crkp_ho_bc) %>% arrange(AICc) %>% select(.model:BIC)

arima_crkp_ho_bc  %>% select(arima_crkp_ho_bc_011_100) %>% gg_tsresiduals() # compare models
arima_crkp_ho_bc %>% select(stepwise_crkp_ho_bc) %>% gg_tsresiduals() # compare models

#### Check portmanteau tests of best fit option
augment(arima_crkp_ho_bc) %>% 
  filter(.model=='arima_crkp_ho_bc_011_100') %>%
  features(.innov, ljung_box, lag = 10, dof = 3)  # portmanteau test on best manual



```

### 3.1.3 HCA-CRAb clinical infections (*crab hca clinical*)
```{r message=FALSE, warning=FALSE, include=FALSE}

#### Plot data, ACF, PACF
tswide_clean %>% gg_tsdisplay(crab_ho_clinical, "partial") 

#### Apply a unit root test to check if time-series stationary,  determine number of non-seasonal (ndiffs) and seasonal (nsdiffs) differences required to make stationary, and plot differenced time-series if required.
tswide_clean %>% features(crab_ho_clinical, unitroot_kpss)
tswide_clean %>% features(crab_ho_clinical, unitroot_ndiffs)
tswide_clean %>% features(crab_ho_clinical, unitroot_nsdiffs,.period=12)
tswide_clean %>% gg_tsdisplay(difference(crab_ho_clinical), "partial") 


#### Examine need for a Box-Cox data transformation to reduce heteroskedasticity (i.e variability in the variance of the time-series)
lambda<-guerrero(tswide_clean$crab_ho_clinical, .period=12L)

#### Fit best manual and auto ARIMA on transformed data
arima_crab_ho <- tswide_clean %>% model(
  arima_crab_ho_111 = ARIMA(crab_ho_clinical ~ pdq(1,1,1)),
  arima_crab_ho_011 = ARIMA(crab_ho_clinical ~ pdq(0,1,1)),
  arima_crab_ho_110 = ARIMA(crab_ho_clinical ~ pdq(1,1,0)),
  arima_crab_ho_101 = ARIMA(crab_ho_clinical ~ pdq(1,0,1)),
  stepwise_crab_ho = ARIMA(crab_ho_clinical))

glance(arima_crab_ho) %>% arrange(AICc) %>% select(.model:BIC)

# check the residuals of best models to ensure white noise.
arima_crab_ho  %>% select(arima_crab_ho_011) %>% gg_tsresiduals() 
arima_crab_ho %>% select(arima_crab_ho_101) %>% gg_tsresiduals()

#### Check portmanteau tests of best fit option
augment(arima_crab_ho) %>% 
  filter(.model=='arima_crab_ho_011') %>%
  features(.innov, ljung_box, lag = 10, dof = 3)  # portmanteau test on best manual


```

### 3.1.4 HCA-CRPa clinical infections (*crpa hca clinical*)
```{r message=FALSE, warning=FALSE, include=FALSE}

#### Plot data, ACF, PACF
tswide_clean %>% gg_tsdisplay(crpa_ho_clinical, "partial")

#### Apply a unit root test to check if time-series stationary,  determine number of non-seasonal (ndiffs) and seasonal (nsdiffs) differences required to make stationary, and plot differenced time-series if required.
tswide_clean %>% features(crpa_ho_clinical, unitroot_kpss)
tswide_clean %>% features(crpa_ho_clinical, unitroot_ndiffs) 
tswide_clean %>% features(crpa_ho_clinical, unitroot_nsdiffs,.period=12) 

tswide_clean %>% gg_tsdisplay(difference(crpa_ho_clinical), "partial") 
crpa_ho_clin_diff1 <-  tsibble::difference(tswide_clean$crpa_ho_clinical, lag =1, differences = 1)
tswide_clean %>% gg_tsdisplay(crpa_ho_clin_diff1, "partial")

#### Examine need for a Box-Cox data transformation to reduce heteroskedasticity (i.e variability in the variance of the time-series)
tswide_clean <- tswide_clean %>% dplyr::mutate(crpa_ho_clin_diff1 = replace_na(crpa_ho_clin_diff1, 0))
tswide_clean <- tswide_clean %>% dplyr::mutate(new_crpa_ho_clin_diff1 = crpa_ho_clin_diff1+10)
boxcoxnc(tswide_clean$new_crpa_ho_clin_diff1, method = "sw", lambda2 = 10)


#### Repeat unit root tests and differencing number evaluations on the transformed data
tswide_clean %>% features(crpa_ho_clin_diff1, unitroot_kpss)
tswide_clean %>% features(crpa_ho_clin_diff1, unitroot_ndiffs)
tswide_clean %>% features(crpa_ho_clin_diff1, unitroot_nsdiffs, .period=12)

#### Fit best manual and auto ARIMA
arima_crpa_ho <- tswide_clean %>% model(
  arima_crpa_ho_111 = ARIMA(crpa_ho_clinical ~ pdq(1,1,1)),
  arima_crpa_ho_011 = ARIMA(crpa_ho_clinical ~ pdq(0,1,1)),
  arima_crpa_ho_101 = ARIMA(crpa_ho_clinical ~ pdq(1,0,1)),
  stepwise_crpa_ho = ARIMA(crpa_ho_clinical)) 

glance(arima_crpa_ho) %>% arrange(AICc) %>% select(.model:BIC)

#### Examine residuals of ARIMA to check correct identification
arima_crpa_ho  %>% select(arima_crpa_ho_011) %>% gg_tsresiduals()
arima_crpa_ho %>% select(arima_crpa_ho_101) %>% gg_tsresiduals()
arima_crpa_ho %>% select(stepwise_crpa_ho) %>% gg_tsresiduals() 

#### Check portmanteau tests of best fit option
augment(arima_crpa_ho) %>% 
  filter(.model=='arima_crpa_ho_101') %>%
  features(.innov, ljung_box, lag = 10, dof = 3)  # portmanteau test on best manual
```

## 3.2 Antibiotic and IPC time series
```{r}

# Nb: we use the pre-intervention dataset "ts_p2_pre_icarba" for fitting the baseline ARIMA with exogenous variables (ARIMAX) for carbapenems. For other antibiotics we should use the  "ts_p3_pre_iabs_iunivscr" data
pre_icarba <- ts_p2_pre_icarba %>% as_tsibble(index="date")
pre_iabs <- ts_p3_pre_iabs_iunivscr %>% as_tsibble(index="date")



```
## 3.3 Influential infection time-series
These time-series are expected to be related to antibiotic use in the population. Bidirectional relationships may exist but we hypothesised that hospital-wide use of important antibiotic exposures driving carbapenem resistance may be responsive to the following:

* All gram-negative bloodstream isolates (BSIs) (*gmneg_all_bsi*)
* Carbapenem and cephalosporin sensitive E.coli BSI (*ecoliccs_all_bsi*)
* 3rd generation cephalosporins resistant E.coli BSI (*ecoli3gcr_all_bsi*)

### 3.3.1 Gram-Neg BSIs
```{r}
### Carbapenems######################

# Prepare differenced time-series
pre_icarba$gmneg_diff <-  tsibble::difference(pre_icarba$gmneg_all_bsi, lag =1, differences = 1)
pre_icarba$carbapenem_diff <-  tsibble::difference(pre_icarba$carbapenems, lag =1, differences = 1)
pre_icarba_ed2 <- pre_icarba [2:72,] %>% as_tsibble(index="date")
pre_icarba_ed4 <- pre_icarba [4:72,] %>% as_tsibble(index="date")

# Explore cross-correlations with raw data
CCF(pre_icarba, carbapenems, gmneg_all_bsi, lag_max=24) %>% autoplot(level=95)

# Explore cross-correlations with differenced data
CCF(pre_icarba_ed2, carbapenem_diff, gmneg_diff, lag_max=24)%>% autoplot(level=95)

# Identify an ARIMAX model for carbapenems 
carba_pre1 <- pre_icarba %>% model(ARIMA(carbapenems  ~ -1 + pdq(0,1,1)+PDQ(0,0,0)+lag(gmneg_all_bsi,2)))
report(carba_pre1)
carba_pre1 %>%  gg_tsresiduals() 

carba_pre2 <- pre_icarba_ed2 %>% model(ARIMA(carbapenems  ~ -1 + pdq(0,1,1)+PDQ(0,0,0)+lag(gmneg_diff,2)))
report(carba_pre2)
carba_pre2 %>%  gg_tsresiduals() 

# Looking at the correlation between gm neg BSI and differenced gm neg BSI (y-axis)
gmneg_bsi<- pre_icarba$gmneg_all_bsi[1:71]
differenced_gm_negbsi <- pre_icarba$gmneg_diff [2:72]
gmneg <- cbind(gmneg_bsi,differenced_gm_negbsi )
gmneg <-gmneg %>% as_tibble()
cor(gmneg$gmneg_bsi, gmneg$differenced_gm_negbsi)
ggplot(gmneg, aes(gmneg_bsi,differenced_gm_negbsi))+geom_point()

```
# 4. Interactions and VARs

# 5. GAMs

# 6. MARS



```{r}
plot(cars)
```

