---
title: "Modena Time Series Analysis"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  pdf_document:
    toc: yes
---

# Introduction
This R notebook provides code to reproduce results for the paper *Effects of antimicrobial stewardship and infection prevention and control on the incidence of carbapenem resistant gram-negative infections in Modena, Italy: a non-linear time-series analysis*. 
It requires installing of necessary packages


Data and code are available at the Github repository
[via this link](https://github.com/TimLawes/modena_carbapenem_resistance)

# 1. Preliminaries
## 1.1 Install copies of required packages to private library
```{r message=FALSE, warning=FALSE, include=FALSE}

# Install renv for private package library if not available
if(!require(renv)){
    install.packages("renv")
}

Sys.setenv(LOCALAPPDATA=renv::paths$library())
renv::restore()
renv::activate()

```

## 1.2 Load required packages and customised plot theme
```{r, messages =FALSE, warning=FALSE}

library(tidyverse) # to enable tidyverse functions
library(earth) # for MARS
library(tseries) # for working with TS data
library(tsutils) #
library(forecast) # for forcasting
library(fpp3) # enables tsibble, and fable packages
library(plotmo) # plotting MARS model from earth
library(mgcv) # for GLM and GAM models
library(smooth)
library(lmtest) # comparing models 
library(stats)
library(grid)
library(lattice)
library(viridis) # for viridis colour palettes
library(hrbrthemes) #for plot themes
library(sysfonts) # for using google fonts
library(lubridate) # for working with datetime variables
library(car) # for Cook's distances
library(strucchange) # for detecting structural changes in time-series
library(ganttrify) # for time lines (Gantt type)
library(patchwork) # for combining different plot types
library(missForest) # for missing value imputation by random forest
library(mgcViz) # for plots and model checks of GAM
source("cochrane_plot_theme.R") # imports a custom ggplot theme defined in other script
source("OutlierTreatment2.R") # imports custom function for detecting extreme outliers and replacing.
```

## 1.3 Load data and process
```{r, messages =FALSE, warning=FALSE, results='hide'}

# Read in the main data 
main <- read_csv ("modena_carbapenem.csv", show_col_types =FALSE)
date <- tibble(date = as.Date("2008-01-01")+0:131)
main2 <- cbind(date,main) %>% dplyr::select(!c(months, study_m))

# Create wide format without time-series index (not as a tsibble)
wide <- main2  %>% mutate(across(gmneg_all_bsi:crpa_cohca_swab,~./obd*10000),
                         across(abhr: fluoroquinolones, ~./obd*1000)) 

# Create wide as a time-series indexed table (tsibble)
tswide <- wide %>% as_tsibble(index = date)

```

## 1.4 Outlier detection and time-series cleaning
```{r, messages=FALSE, warning=FALSE}

# outlier detection and replacement with cleaned values
wideabx <- wide %>% as_tibble %>% dplyr::select(date, obd, bedocc,
abhr:fluoroquinolones)

newabx.matrix<-rep(0,132)
for(i in 1:length(wideabx[1,])){
abx<- wideabx[,i]
newabx<-OutlierTreatment2(abx,freq=12,start.date=c(2008,1), only.clean=TRUE)
newabx.matrix <-cbind(newabx.matrix, newabx)
}
newabx.matrix<-newabx.matrix[,-1]

colnames(newabx.matrix)<-c(paste0(colnames(wideabx)))
wide_clean <-main2 %>% dplyr::select(!c(date, obd,bedocc,
abhr:fluoroquinolones))
wide_clean <- cbind(wide_clean, newabx.matrix)

```


## 1.5 Trend decomposition
```{r, messages =FALSE, warning=FALSE}

# create long format data for incidence

tslong <- tswide %>% dplyr::select(!c(hh_adherence:mrsa_all_bsi)) %>%
  pivot_longer(
  cols = csab_all_any:crpa_cohca_swab,
  names_to = c("IR", "org", "pop", "site"),
  names_pattern = "(cr|cs)(ab|kp|pa)_(.*)_(.*)") %>% filter(pop!="NA") 
tslong_incid <- tslong %>% mutate(incidence = value/obd*10000)
tslong_incid <- tslong_incid %>% mutate (bugs = paste(IR,org,pop,site)) %>% dplyr::select(-"value")

tslong_bugs <- tslong_incid %>% dplyr::select(-bedocc) %>% group_by(bugs,IR,org,pop,site) %>% nest() 

tslongbugs2<- tslong_bugs %>% filter(pop=="hca" | pop=="conothca")

decompose <- function(data, y, ...){
    y <- enquo(y)
    dvar <- data %>% pull(!!y) %>% as.ts()
    tmpd <- data %>% model(STL(dvar ~ trend(window = 7) + 
                                 season(window =13), robust = TRUE)) %>%
      components()
    data <-cbind(data,tmpd[,3:6])
}

ts_decomp <- tslongbugs2 %>% 
    mutate(decs = map(data, decompose, y = incidence)) %>%
    dplyr::select(-data) %>% unnest(decs) %>% ungroup()
ts_decomp <- ts_decomp %>% mutate (trend=if_else(trend <0, 0, trend))

```

# 2. Descriptives

## 2.1 Regional & national comparison
```{r, messages =FALSE, warning=FALSE}

# Prepare data
otherpops <- read_csv("populations_year.csv", col_types = cols(
  pop = col_factor(),
  org = col_factor(),
  poporg = col_factor(),
  orgsite = col_factor(),
  year = col_integer(),
  site = col_factor(),
  num_ts=col_factor()
))

otherpops_ed <- otherpops %>% filter(studyp==1) %>% mutate(y=pc_IR/100)
otherpops_ed <- otherpops_ed %>% filter(carb_IR != "NA") %>% mutate(pop_org = interaction(pop, org))
otherpops_ed1 <- otherpops_ed %>% mutate(
  opop = ordered(pop, levels = c("Emilia-Romagna (other)","Emilia-Romagna",  "Modena", "Italy")),
  oorg = ordered(org, levels=c("P.aeruginosa","A.baumannii", "K.pneumoniae")),
  osite = ordered(site,levels = c("All clinical", "BSI")))
otherpops_ed2 <- otherpops_ed %>% filter(pop=="Modena")


#### Observed value plot ####################
# Plot of % resistant (%I/R) by population and organism
popp1<- ggplot(otherpops_ed1,  
    aes(year,pc_IR, group=pop, color=pop))+
    geom_line()+
    scale_x_continuous("Year",breaks=seq(2008,2018,2))+
    facet_grid(org~site, scales = "free_y")+
    scale_color_viridis_d(option="A", begin=0.8, end=0.2)+
    theme_pnaat()+
  theme(legend.position="top")

# Plot the % of regional infections in Modena by carbapenem sensitivity
popp2<- ggplot(otherpops_ed2, aes(year,`pc_Emilia-Romagna_IR`, 
  group= org, color=org))+
  geom_line()+
  geom_line(aes(year,`pc_Emilia-Romagna_S`), linetype = "dashed")+
  scale_x_continuous("Year", breaks=seq(2008,2018,2))+
  facet_grid(org~site, scales="free_y")+
  scale_color_viridis_d(option="A", begin=0.8, end=0.2)+
  theme_pnaat()+
  theme(legend.position="top")

popp1+popp2
  

###### GAMs by organism ############################################

### K.pneumoniae GAM
kp_pop <- otherpops_ed1 %>% filter(org =="K.pneumoniae")

# "ordered factor" GAM to identify between population differences
kp_pop_GAM <- gam(carb_IR/tested ~ osite+opop +s(year) +s(year,by=opop, m=2, k=-1, bs="tp")+s(year,by=osite, k=-1, m=2, bs="tp"), data = kp_pop, weights = tested, family = binomial(link="logit"))

acf_resid(kp_pop_GAM , split_pred=list(pop= kp_pop$opop, site=kp_pop$site),n=6)
(valRho <- acf(resid(kp_pop_GAM), plot=FALSE)$acf[2])

kppopViz <- getViz(kp_pop_GAM)
summary(kppopViz)
gam.check(kppopViz)
print(plot(kppopViz, allTerms = T), pages = 1)
concurvity(kppopViz, full=TRUE)
AIC(kppopViz)
BIC(kppopViz)

# With random smooths
kp_pop_GAMrand <- gam(carb_IR/tested ~ s(year, pop, bs="fs")+s(year,site, bs="fs"), data = kp_pop, weights = tested, family = binomial(link="logit"))

acf_resid(kp_pop_GAMrand, split_pred=list(pop= kp_pop$opop, site=kp_pop$site),n=6)
(valRho <- acf(resid(kp_pop_GAMrand), plot=FALSE)$acf[2])

kppopViz1 <- getViz(kp_pop_GAMrand)
summary(kppopViz1)
gam.check(kppopViz1)
print(plot(kppopViz1, allTerms = T), pages = 1)
concurvity(kppopViz1, full=TRUE)
AIC(kppopViz1)
BIC(kppopViz1)

# Smooth by Interaction of factors
kp_pop_GAM2 <- gam(carb_IR/tested ~ opop+osite+ s(year,by=interaction(osite,opop), k=-1), data =  kp_pop, weights = tested, family = binomial(link="logit"))

popViz <- getViz(kp_pop_GAM2 )
summary(popViz)
check(popViz)
print(plot(popViz, allTerms = T), pages = 1)
concurvity(popViz, full=TRUE)
AIC(popViz)
BIC(popViz)

# gather and plot
kp_pop_fit<- cbind(kp_pop, predict.gam(kp_pop_GAM2,
                  se.fit=TRUE, type = "link")) 
kp_pop_fit$uci<-kp_pop_fit$fit + (2 * kp_pop_fit$se.fit) 
kp_pop_fit$lci<-kp_pop_fit$fit - (2 * kp_pop_fit$se.fit) 
kp_pop_fit$uci<-kp_pop_GAM$family$linkinv(kp_pop_fit$uci )
kp_pop_fit$lci<-kp_pop_GAM$family$linkinv(kp_pop_fit$lci )
kp_pop_fit$est<- kp_pop_GAM$family$linkinv(kp_pop_fit$fit)

kppop_plot <- ggplot(kp_pop_fit, aes(x=year, y= fit, color=pop, fill=pop))+
  geom_point(aes(x=year, y=(carb_IR/tested)))+
  geom_line(aes(x=year, y=est))+
  geom_ribbon(aes(x=year, ymin=lci, ymax=uci), alpha=0.3, color=NA)+facet_grid(~site)+
  scale_x_continuous(name="Year", breaks=seq(2008,2018,2))+
  scale_y_continuous(name="%I/R", labels=scales::percent_format(accuracy=1))+
  scale_color_viridis_d(option="A", begin = 0.2, end=0.8)+
  scale_fill_viridis_d(option="A", begin = 0.2, end=0.8)

##### P.aeruginosa GAM
pa_pop <- otherpops_ed1 %>% filter(org =="P.aeruginosa")

# "ordered factor" GAM to identify between population differences
pa_pop_GAM <- gam(carb_IR/tested ~ opop*osite+s(year)+s(year, by=opop, m=2, bs="tp")+s(year, by=osite, m=2, bs="tp"), data = pa_pop, weights = tested, family = binomial(link="logit"))

acf_resid(pa_pop_GAM, split_pred=list(pop= pa_pop$opop, site=pa_pop$osite),n=6)
(valRho <- acf(resid(pa_pop_GAM), plot=FALSE)$acf[2])

papopViz <- getViz(pa_pop_GAM)
summary(papopViz)
check(papopViz)
print(plot(papopViz, allTerms = T), pages = 1)
concurvity(papopViz, full=TRUE)
concurvity(papopViz, full=FALSE)
AIC(papopViz)
BIC(papopViz)

# With random smooths
pa_pop_GAMrand <- gam(carb_IR/tested ~ s(year, by=pop, bs="fs")+s(site, bs="re"), data = pa_pop, weights = tested, family = binomial(link="logit"))

acf_resid(pa_pop_GAMrand, split_pred=list(pop= pa_pop$opop, site=pa_pop$site),n=6)
(valRho <- acf(resid(pa_pop_GAMrand), plot=FALSE)$acf[2])

papopViz1 <- getViz(pa_pop_GAMrand)
summary(papopViz1)
gam.check(papopViz1)
print(plot(papopViz1, allTerms = T), pages = 1)
concurvity(papopViz1, full=TRUE)
AIC(papopViz1)
BIC(papopViz1)

# Smooth by Interaction of factors
pa_pop_GAM2 <- gam(carb_IR/tested ~ opop+osite+ s(year,by=interaction(osite,opop), k=-1), data =  pa_pop, weights = tested, family = binomial(link="logit"))

popViz <- getViz(kp_pop_GAM2 )
summary(popViz)
check(popViz)
print(plot(popViz, allTerms = T), pages = 1)
concurvity(popViz, full=TRUE)
AIC(popViz)
BIC(popViz)

# gather and plot predicted values
pa_pop_fit<- cbind(pa_pop, predict.gam(pa_pop_GAM2,
                  se.fit=TRUE, type = "link")) 
pa_pop_fit$uci<-pa_pop_fit$fit + (2 * pa_pop_fit$se.fit) 
pa_pop_fit$lci<-pa_pop_fit$fit - (2 * pa_pop_fit$se.fit) 
pa_pop_fit$uci<-pa_pop_GAM$family$linkinv(pa_pop_fit$uci )
pa_pop_fit$lci<-pa_pop_GAM$family$linkinv(pa_pop_fit$lci )
pa_pop_fit$est<- pa_pop_GAM$family$linkinv(pa_pop_fit$fit)

papop_plot <- ggplot(pa_pop_fit, aes(x=year, y= fit, color=pop, fill=pop))+
  geom_point(aes(x=year, y=(carb_IR/tested)))+
  geom_line(aes(x=year, y=est))+
  geom_ribbon(aes(x=year, ymin=lci, ymax=uci), alpha=0.3, color=NA)+facet_grid(~site)+
  scale_x_continuous(name="Year", breaks=seq(2008,2018,2))+
  scale_y_continuous(name="%I/R", labels=scales::percent_format(accuracy=1))+
  scale_color_viridis_d(option="A", begin = 0.2, end=0.8)+
  scale_fill_viridis_d(option="A", begin = 0.2, end=0.8)

##### A.baumannii GAM
ab_pop <- otherpops_ed1 %>% filter(org =="A.baumannii")

# "ordered factor" GAM to identify between population differences
ab_pop_GAM <- gam(carb_IR/tested ~ opop+osite+ s(year)+s(year, by=opop, m=2, bs="fs")+s(year, by=osite, m=2, bs="fs"), data = ab_pop, weights = tested, family = binomial(link="logit"))

abpopViz <- getViz(ab_pop_GAM)
summary(abpopViz)
check(abpopViz)
print(plot(abpopViz, allTerms = T),pages = 1)
concurvity(abpopViz, full=TRUE)
concurvity(abpopViz, full=FALSE)
AIC(abpopViz)
BIC(abpopViz)

acf_resid(ab_pop_GAM, split_pred=list(pop= ab_pop$opop, site=ab_pop$site),n=6)
(valRho <- acf(resid(ab_pop_GAM), plot=FALSE)$acf[2])


# Random smooths GAM 
ab_pop_GAM1 <- gam(carb_IR/tested ~ s(year, pop, m=2, bs="fs"), data = ab_pop, weights = tested, family = binomial(link="logit"))

abpopViz <- getViz(ab_pop_GAM1)
summary(abpopViz)
check(abpopViz)
print(plot(abpopViz, allTerms = T),pages = 1)
concurvity(abpopViz, full=TRUE)
concurvity(abpopViz, full=FALSE)
AIC(abpopViz)
BIC(abpopViz)

acf_resid(ab_pop_GAM1, split_pred=list(pop= ab_pop$opop, site=ab_pop$site),n=6)
(valRho <- acf(resid(ab_pop_GAM1), plot=FALSE)$acf[2])

# Smooth by Interaction of factors
ab_pop_GAM2 <- gam(carb_IR/tested ~ opop+osite+ s(year,by=interaction(osite,opop), k=-1), data =  ab_pop, weights = tested, family = binomial(link="logit"))

popViz <- getViz(ab_pop_GAM2 )
summary(popViz)
check(popViz)
print(plot(popViz, allTerms = T), pages = 1)
concurvity(popViz, full=TRUE)
AIC(popViz)
BIC(popViz)

# gather and plot predicted values
ab_pop_fit<- cbind(ab_pop, predict.gam(ab_pop_GAM2,
                  se.fit=TRUE, type = "link")) 
ab_pop_fit$uci<-ab_pop_fit$fit + (2 * ab_pop_fit$se.fit) 
ab_pop_fit$lci<-ab_pop_fit$fit - (2 * ab_pop_fit$se.fit) 
ab_pop_fit$uci<-ab_pop_GAM$family$linkinv(ab_pop_fit$uci )
ab_pop_fit$lci<-ab_pop_GAM$family$linkinv(ab_pop_fit$lci )
ab_pop_fit$est<- ab_pop_GAM$family$linkinv(ab_pop_fit$fit)

abpop_plot <- ggplot(ab_pop_fit, aes(x=year, y= fit, color=pop, fill=pop))+
  geom_point(aes(x=year, y=(carb_IR/tested)))+
  geom_line(aes(x=year, y=est))+
  geom_ribbon(aes(x=year, ymin=lci, ymax=uci), alpha=0.3, color=NA)+facet_grid(~site)+
  scale_x_continuous(name="Year", breaks=seq(2008,2018,2))+
  scale_y_continuous(name="%I/R", labels=scales::percent_format(accuracy=1))+
  scale_color_viridis_d(option="A", begin = 0.2, end=0.8)+
  scale_fill_viridis_d(option="A", begin = 0.2, end=0.8)


################### Plots by site of infection #################

#### Clinical GAM
clin_pop <- otherpops_ed1 %>% filter(site=="All clinical")

# "ordered factor" GAM to identify between population differences
clinical_GAM <- gam(carb_IR/tested ~ opop*oorg +s(year)+s(year, by=opop,m=2, bs="tp", k=-1)+s(year, by=oorg,m=2, k=-1), data =  clin_pop, weights = tested, family = binomial(link="logit"), method="REML")
                    
popViz <- getViz(clinical_GAM)
summary(popViz)
check(popViz)
print(plot(popViz, allTerms = T), pages = 1)
concurvity(popViz, full=TRUE)
concurvity(popViz, full=FALSE)
AIC(popViz)
BIC(popViz)

acf_resid(clinical_GAM, split_pred=list(pop= clin_pop$opop, site=clin_pop$osite),n=6)
(valRho <- acf(resid(clinical_GAM), plot=FALSE)$acf[2])

# random smooths model
clinical_GAM1<- gam(carb_IR/tested ~ s(year, pop,k=-1, bs="fs")+s(year, org,k=-1, bs="fs"), data =  clin_pop, weights = tested, family = binomial(link="logit"), AR.start= clin_pop$start.event, rho=0.9)

popViz <- getViz(clinical_GAM1)
summary(popViz)
check(popViz)
print(plot(popViz, allTerms = T), pages = 1)
concurvity(popViz, full=TRUE)
AIC(popViz)
BIC(popViz)

acf_resid(clinical_GAM1, split_pred=list(pop= clin_pop$opop, site=clin_pop$osite),n=6)
(valRho <- acf(resid(clinical_GAM2), plot=FALSE)$acf[2])

# Smooth by Interaction of factors
clinical_GAM2 <- gam(carb_IR/tested ~ opop*oorg+ s(year,by=interaction(oorg,opop), k=-1), data =  clin_pop, weights = tested, family = binomial(link="logit"))

popViz <- getViz(clinical_GAM2)
summary(popViz)
check(popViz)
print(plot(popViz, allTerms = T), pages = 1)
concurvity(popViz, full=TRUE)
AIC(popViz)
BIC(popViz)


# gather and plot predicted values
clin_fit<- cbind(clin_pop, predict.gam(clinical_GAM2,
                  se.fit=TRUE, type = "link")) 
clin_fit$uci<-clin_fit$fit + (2 * clin_fit$se.fit) 
clin_fit$lci<-clin_fit$fit - (2 * clin_fit$se.fit) 
clin_fit$uci<-clinical_GAM$family$linkinv(clin_fit$uci )
clin_fit$lci<-clinical_GAM$family$linkinv(clin_fit$lci )
clin_fit$est<- clinical_GAM$family$linkinv(clin_fit$fit)

clin_pop_plot<- ggplot(clin_fit, aes(x=year, y= est, color=pop, fill=pop))+
  geom_point(aes(x=year, y=(carb_IR/tested)), size=1)+
  geom_line(aes(x=year, y=est))+
  geom_ribbon(aes(x=year, ymin=lci, ymax=uci), alpha=0.2, color=NA)+
  facet_grid(site~org)+
  scale_x_continuous(name="Year", breaks=seq(2008,2018,2))+
  scale_y_continuous(name="%I/R", labels=scales::percent_format(accuracy=1))+
  theme_pnaat()+
  scale_color_viridis_d(option="A", begin = 0.2, end=0.8)+
  scale_fill_viridis_d(option="A", begin = 0.2, end=0.8)


#### BSI GAM
bsi_pop <- otherpops_ed1 %>% filter(site=="BSI")

# "ordered factor" GAM to identify between population differences
bsi_GAM <- gam(carb_IR/tested ~ opop*oorg+s(year)+s(year,by=oorg, k=-1, bs="tp",m=2)+s(year,by=opop, k=-1, bs="tp",m=2), data =  bsi_pop, weights = tested, family = binomial(link="logit"))

popViz <- getViz(bsi_GAM)
summary(popViz)
check(popViz)
print(plot(popViz, allTerms = T), pages = 1)
concurvity(popViz, full=TRUE)
concurvity(popViz, full=FALSE)
AIC(popViz)
BIC(popViz)

acf_resid(bsi_GAM, split_pred=list(pop= bsi_pop$opop, site=bsi_pop$osite),n=6)
(valRho <- acf(resid(bsi_GAM), plot=FALSE)$acf[2])

# Random smooths approach
bsi_GAM1 <- gam(carb_IR/tested ~ s(year, pop, k=-1, bs="fs",m=1)+ s(year, org, k=-1, bs="fs",m=1), data =  bsi_pop, weights = tested, family = binomial(link="logit"))

popViz <- getViz(bsi_GAM1)
summary(popViz)
check(popViz)
print(plot(popViz, allTerms = T), pages = 1)
concurvity(popViz, full=TRUE)
concurvity(popViz, full=FALSE)
AIC(popViz)
BIC(popViz)

acf_resid(bsi_GAM1, split_pred=list(pop= bsi_pop$opop, site=bsi_pop$osite),n=6)
(valRho <- acf(resid(bsi_GAM2), plot=FALSE)$acf[2])


# Smooth by Interaction of factors
BSI_GAM2 <- gam(carb_IR/tested ~ opop*oorg+ s(year,by=interaction(oorg,opop), k=-1), data =  clin_pop, weights = tested, family = binomial(link="logit"))

popViz <- getViz(BSI_GAM2)
summary(popViz)
check(popViz)
print(plot(popViz, allTerms = T), pages = 1)
concurvity(popViz, full=TRUE)
AIC(popViz)
BIC(popViz)

# gather and plot predicted values
bsi_fit<- cbind(bsi_pop, predict.gam(bsi_GAM2,
                  se.fit=TRUE, type = "link")) 
bsi_fit$uci<-bsi_fit$fit + (2 * bsi_fit$se.fit) 
bsi_fit$lci<-bsi_fit$fit - (2 * bsi_fit$se.fit) 
bsi_fit$uci<-bsi_GAM$family$linkinv(bsi_fit$uci )
bsi_fit$lci<-bsi_GAM$family$linkinv(bsi_fit$lci )
bsi_fit$est<- bsi_GAM$family$linkinv(bsi_fit$fit)

bsi_pop_plot<- ggplot(bsi_fit, aes(x=year, y= est, color=pop, fill=pop))+
  geom_point(aes(x=year, y=(carb_IR/tested)))+
  geom_line(aes(x=year, y=est))+
  geom_ribbon(aes(x=year, ymin=lci, ymax=uci), alpha=0.2, color=NA)+
  facet_grid(site~org)+
  scale_x_continuous(name="Year", breaks=seq(2008,2018,2))+
  scale_y_continuous(name="%I/R", labels=scales::percent_format(accuracy=1))+
  scale_color_viridis_d(option="A", begin = 0.2, end=0.8)+
  scale_fill_viridis_d(option="A", begin = 0.2, end=0.8)

##### Summary plots #########
bysite_pop_fit<-rbind(bsi_fit, clin_fit)

bysite_pop_plot <- ggplot(bysite_pop_fit, aes(x=year, y= fit, color=pop, fill=pop))+
  geom_point(aes(x=year, y=(carb_IR/tested)), size=1)+
  geom_line(aes(x=year, y=est))+
  geom_ribbon(aes(x=year, ymin=lci, ymax=uci), alpha=0.1,
              color=NA)+facet_grid(org~site, scales="free_y")+
  scale_x_continuous(name="Year", breaks=seq(2008,2018,2))+
  scale_y_continuous(name="%I/R", labels=scales::percent_format(accuracy=1))+
  theme_pnaat()+
  scale_color_viridis_d(option="A", begin = 0.2, end=0.8)+
  scale_fill_viridis_d(option="A", begin = 0.2, end=0.8)

byorg_pop_fit<-rbind(kp_pop_fit, pa_pop_fit, ab_pop_fit)

byorg_pop_plot <- ggplot(byorg_pop_fit, aes(x=year, y= fit, color=pop, fill=pop))+
  geom_point(aes(x=year, y=(carb_IR/tested)), size=1)+
  geom_line(aes(x=year, y=est))+
  geom_ribbon(aes(x=year, ymin=lci, ymax=uci), alpha=0.1,
              color=NA)+facet_grid(org~site, scales="free_y")+
  scale_x_continuous(name="Year", breaks=seq(2008,2018,2))+
  scale_y_continuous(name="%I/R", labels=scales::percent_format(accuracy=1))+
  theme_pnaat()+
  scale_color_viridis_d(option="A", begin = 0.2, end=0.8)+
  scale_fill_viridis_d(option="A", begin = 0.2, end=0.8)

byorg_pop_fit_ed <- byorg_pop_fit %>% filter(pop =="Modena"| pop =="Emilia-Romagna (other)")

byorg_pop_plot2 <- ggplot(byorg_pop_fit_ed, aes(x=year, y= fit, color=site, fill=site))+
  geom_line(aes(x=year, y=est))+
  geom_ribbon(aes(x=year, ymin=lci, ymax=uci), alpha=0.2,
              color=NA)+facet_grid(org~pop, scales="free_y")+
  scale_x_continuous(name="Year", breaks=seq(2008,2018,2))+
  scale_y_continuous(name="%I/R", labels=scales::percent_format(accuracy=1))+
  theme_pnaat()+
  scale_color_viridis_d(option="A", begin = 0.2, end=0.8)+
  scale_fill_viridis_d(option="A", begin = 0.2, end=0.8)

byorg_pop_plot
bysite_pop_plot
byorg_pop_plot2
```

### 2.1.2 Quantify differences
```{r, messages =FALSE, warning=FALSE}

## Were "BSI" trends significantly different from clinical?

# Diffs for K.pneumoniae
kpdiff_modena <- plot_diff(kp_pop_GAM2, view="year", comp=list(osite=c('BSI', 'All clinical')), cond=list(opop="Modena"), n.grid=1000, plot_all="site", rm.ranef=FALSE, rug=FALSE, sim.ci=TRUE, plot=FALSE)

kp_mod_clinical <- get_predictions(kp_pop_GAM2,
cond=list(opop ='Modena', osite= "All clinical", year=seq(2008,2018,length=1000)),
rm.ranef=FALSE, sim.ci=TRUE)

kp_diff_mod <- left_join(kp_mod_clinical, kpdiff_modena, by="year")

kp_diff_mod <-kp_diff_mod %>% mutate(
  differences = plogis(fit+est)-plogis(fit),
  lci = plogis(fit+est-sim.CI.y)-plogis(fit),
  uci = plogis(fit+est+sim.CI.y)-plogis(fit),
  clinical_fit = plogis(fit),
  clinical_lci = plogis(fit-sim.CI.x),
  clinical_uci = plogis(fit+sim.CI.x),
  BSI_fit = plogis(fit+est),
  BSI_lci = plogis((fit+est)-sim.CI.y),
  BSI_uci = plogis((fit+est)+sim.CI.y),
  mean_fit = (clinical_fit+BSI_fit)/2,
  organism = "K.pneumoniae",
  partial_diffs = plogis(fit+(est+0.44035))-plogis(fit)
)

# Diffs for P.aeurginosa
padiff_modena <- plot_diff(pa_pop_GAM2, view="year", comp=list(osite=c('BSI', 'All clinical')), cond=list(opop="Modena"), n.grid=1000, plot_all="site", rm.ranef=FALSE, rug=FALSE, sim.ci=TRUE, plot=FALSE)

pa_mod_clinical <- get_predictions(pa_pop_GAM2,
cond=list(opop ='Modena', osite= "All clinical", year=seq(2008,2018,length=1000)),
rm.ranef=FALSE, sim.ci=TRUE)

pa_diff_mod <- left_join(pa_mod_clinical, padiff_modena, by="year")

pa_diff_mod <-pa_diff_mod %>% mutate(
  differences = plogis(fit+est)-plogis(fit),
  lci = plogis(fit+est-sim.CI.y)-plogis(fit),
  uci = plogis(fit+est+sim.CI.y)-plogis(fit),
  clinical_fit = plogis(fit),
  clinical_lci = plogis(fit-sim.CI.x),
  clinical_uci = plogis(fit+sim.CI.x),
  BSI_fit = plogis(fit+est),
  BSI_lci = plogis((fit+est)-sim.CI.y),
  BSI_uci = plogis((fit+est)+sim.CI.y),
  mean_fit = (clinical_fit+BSI_fit)/2,
  organism = "P.aeruginosa",
  partial_diffs = plogis(fit+est-0.04075)-plogis(fit)
)

# Diffs for A.baumannii
abdiff_modena <- plot_diff(ab_pop_GAM2, view="year", comp=list(osite=c('BSI', 'All clinical')), cond=list(opop="Modena"), n.grid=1000, plot_all="site", rm.ranef=FALSE, rug=FALSE, sim.ci=TRUE, plot=FALSE)

ab_mod_clinical <- get_predictions(ab_pop_GAM2,
cond=list(opop ='Modena', osite= "All clinical", year=seq(2008,2018,length=1000)),
rm.ranef=FALSE, sim.ci=TRUE)

ab_diff_mod <- left_join(ab_mod_clinical, abdiff_modena, by="year")

ab_diff_mod <-ab_diff_mod %>% mutate(
  differences = plogis(fit+est)-plogis(fit),
  lci = plogis(fit+est-sim.CI.y)-plogis(fit),
  uci = plogis(fit+est+sim.CI.y)-plogis(fit),
  clinical_fit = plogis(fit),
  clinical_lci = plogis(fit-sim.CI.x),
  clinical_uci = plogis(fit+sim.CI.x),
  BSI_fit = plogis(fit+est),
  BSI_lci = plogis((fit+est)-sim.CI.y),
  BSI_uci = plogis((fit+est)+sim.CI.y),
  mean_fit = (clinical_fit+BSI_fit)/2,
  organism = "A.baumannii",
  partial_diffs = plogis(fit+est+0.13839)
)

# Final plots
diffmods <- rbind(pa_diff_mod, kp_diff_mod, ab_diff_mod)
diffmods_long <- diffmods %>% pivot_longer(cols=c(BSI_fit,BSI_lci,BSI_uci, clinical_fit, clinical_lci, clinical_uci), names_to=c("site",".value"),
                                           names_pattern="(.)_(.*)")

modena_compare<- ggplot(data=diffmods, aes(x=year,y=clinical_fit))+
  geom_line(color="#fd9668")+
  geom_hline(yintercept=0, linetype="dashed", color="grey")+
  geom_ribbon(aes(ymin=(clinical_lci),
                  ymax=(clinical_uci)), 
              fill="#fd9668", alpha=0.1)+
  geom_line(aes(x=year, y=BSI_fit), color="#440f76")+
  geom_ribbon(aes(ymin=(BSI_lci), ymax=(BSI_uci)), 
              alpha=0.1, fill="#440f76")+
  scale_x_continuous(name="Year", breaks=seq(2008,2018,2))+
  scale_y_continuous(name = "%I/R", breaks=seq(0,1,0.1),
                     labels=scales::percent_format(accuracy=1))+
  facet_grid(organism~., scales="free_y")+
  theme_pnaat()

modena_totaldiffs<- ggplot(data=diffmods, aes(x=year,y=differences))+
  geom_line(color="#cd4071")+
  geom_hline(yintercept=0, linetype="dashed", color="grey")+
  geom_ribbon(aes(ymin=(lci),
                  ymax=(uci)), alpha=0.1, fill="#cd4071")+
  scale_x_continuous(name="Year", breaks=seq(2008,2018,2))+
  scale_y_continuous(name = "Total difference in %I/R (BSI vs. All Clinical)",   breaks=seq(-5,5,0.1), limits=c(-0.15,0.45),
labels=scales::percent_format(accuracy=1))+
  facet_grid(organism~.)+
  theme_pnaat()


gridPrint(
plot(sm(kppopViz,3)) + l_fitLine(colour = "red") +l_rug(mapping = aes(x, y), alpha= 0.8)+ l_ciPoly(mul=2, fill = "blue", alpha = 0.1)+
    l_points(shape = 19, size = 1, alpha = 0.1),
plot(sm(papopViz, 3))+ l_fitLine(colour = "red") +l_rug(mapping = aes(x, y), alpha= 0.8)+l_ciPoly(mul=2, fill = "blue", alpha = 0.1)+
    l_points(shape = 19, size = 1, alpha = 0.1),
plot(sm(abpopViz, 3))+ l_fitLine(colour = "red") +l_rug(mapping = aes(x, y), alpha= 0.8)+l_ciPoly(mul=2, fill = "blue", alpha = 0.1)+
    l_points(shape = 19, size = 1, alpha = 0.1))

plot_diff(kp_pop_GAM2, view="year",         
        comp=list(osite=c("BSI", "All clinical")),
        main='Time difference Adults-Children')
plot_diff(pa_pop_GAM2, view="year",         
        comp=list(osite=c("BSI", "All clinical")),
        main='Time difference Adults-Children')
plot_diff(ab_pop_GAM2, view="year",         
        comp=list(osite=c("BSI", "All clinical")),
        main=NULL)


modena_compare + modena_totaldiffs


partial_kp2 + partial_pa2 +partial_ab2

gridPrint(ck1, ck2, ncol = 2)
gridPrint(plot(sm(b, 1)) + l_fitRaster() + l_fitContour() + labs(title = NULL) + guides(fill=FALSE),
          plot(pterm(b, 1)) + l_ciPoly() + l_fitLine(), ncol = 2)
```


# 3. ARIMAs
In addition to time-series decomposition applied to all variables (to facilitate descriptive statistics) we further characterised key time-series, including:

* Primary outcomes
* Antibiotic time-series - for counterfactuals
* Other infection time-series driving antibiotic use


## 3.1 Primary outcomes
The primary study outcomes of interest are:

* crpa hca clinical, 
* crpa conothca clinical, 
* crkp hca clinical, 
* crkp conothca clinical, 
* crab hca clinical, 
* crab conothca clinical

### 3.1.1 Redefine timeseries datasets for ARIMAs
```{r message=FALSE, warning=FALSE, include=FALSE}

# Redefine clean time-series as a tsibble
tswide_clean <- wide_clean %>% mutate(tsmonth = yearmonth(date)) %>% as_tsibble(index = tsmonth)

# Derive pre-intervention time-series
ts_p1_pre_ihh_itargscr <- tswide_clean %>% filter (phase <=1)
ts_p2_pre_icarba <- tswide_clean %>% filter (phase <=2)
ts_p3_pre_iabs_iunivscr <- tswide_clean %>% filter (phase <=3)

##### auto arima function for reference

auto.arima.df <- function(data, y, ...){
    y <- enquo(y)
    yts <- data %>% 
        pull(!!y) %>% 
        as.ts()
    auto.arima(yts, ...)
}

```

### 3.1.2 HCA-CRKp clinical infections (*crkp hca clinical*)
```{r message=FALSE, warning=FALSE, include=FALSE}

#### Plot data, ACF, PACF
tswide_clean %>% gg_tsdisplay(crkp_ho_clinical, "partial") 

#### Apply a unit root test to check if time-series stationary,  determine number of non-seasonal (ndiffs) and seasonal (nsdiffs) differences required to make stationary, and plot differenced time-series if required.
tswide_clean %>% features(crkp_ho_clinical, unitroot_kpss)
tswide_clean %>% features(crkp_ho_clinical, unitroot_ndiffs) 
tswide_clean %>% features(crkp_ho_clinical, unitroot_nsdiffs,.period=12) 

#### Examine need for a Box-Cox data transformation to reduce heteroskedasticity (i.e variability in the variance of the time-series)
lambda<-guerrero(tswide_clean$crkp_ho_clinical, .period=12L)

#### create new transformed variable
tswide_clean$crkp_ho_clinical_bc <- (((tswide$crkp_ho_clinical)^lambda)-1)/lambda

####  Plot the transformed variable, ACF, PACF
tswide_clean %>% gg_tsdisplay(crkp_ho_clinical_bc, "partial")

#### Repeat unit root tests and differencing number evaluations on the transformed data
tswide_clean %>% features(crkp_ho_clinical_bc, unitroot_kpss)
tswide_clean %>% features(crkp_ho_clinical_bc, unitroot_ndiffs)
tswide_clean %>% features(crkp_ho_clinical_bc, unitroot_nsdiffs,.period=12) 

tswide_clean %>% gg_tsdisplay(difference(crkp_ho_clinical_bc), "partial")  

# fit best manual and auto ARIMA with default approach for NON-TRANSFORMED DATA
arima_crkp_ho <- tswide_clean %>% model(
  arima_crkp_ho_101 = ARIMA(crkp_ho_clinical ~ pdq(1,0,1)),
  stepwise_crkp_ho = ARIMA(crkp_ho_clinical)) 

glance(arima_crkp_ho) %>% arrange(AICc) %>% select(.model:BIC)
arima_crkp_ho  %>% select(arima_crkp_ho_101) %>% gg_tsresiduals() # compare models
arima_crkp_ho %>% select(stepwise_crkp_ho) %>% gg_tsresiduals() # compare models


#### Fit best manual and auto ARIMA on transformed data
arima_crkp_ho_bc <- tswide_clean %>% model(
  arima_crkp_ho_bc_111 = ARIMA(crkp_ho_clinical_bc ~ pdq(1,1,1)),
  arima_crkp_ho_bc_101_100 = ARIMA(crkp_ho_clinical_bc ~ pdq(1,0,1) + PDQ(1,0,0)),
    arima_crkp_ho_bc_011_100 = ARIMA(crkp_ho_clinical_bc ~ pdq(0,1,1) + PDQ(1,0,0)),
  arima_crkp_ho_bc_101 = ARIMA(crkp_ho_clinical_bc ~ pdq(1,0,1)),
  arima_crkp_ho_bc_001 = ARIMA(crkp_ho_clinical_bc ~ pdq(0,0,1)),
  stepwise_crkp_ho_bc = ARIMA(crkp_ho_clinical_bc)) 

glance(arima_crkp_ho_bc) %>% arrange(AICc) %>% select(.model:BIC)

arima_crkp_ho_bc  %>% select(arima_crkp_ho_bc_011_100) %>% gg_tsresiduals() # compare models
arima_crkp_ho_bc %>% select(stepwise_crkp_ho_bc) %>% gg_tsresiduals() # compare models

#### Check portmanteau tests of best fit option
augment(arima_crkp_ho_bc) %>% 
  filter(.model=='arima_crkp_ho_bc_011_100') %>%
  features(.innov, ljung_box, lag = 10, dof = 3)  # portmanteau test on best manual



```

### 3.1.3 HCA-CRAb clinical infections (*crab hca clinical*)
```{r message=FALSE, warning=FALSE, include=FALSE}

#### Plot data, ACF, PACF
tswide_clean %>% gg_tsdisplay(crab_ho_clinical, "partial") 

#### Apply a unit root test to check if time-series stationary,  determine number of non-seasonal (ndiffs) and seasonal (nsdiffs) differences required to make stationary, and plot differenced time-series if required.
tswide_clean %>% features(crab_ho_clinical, unitroot_kpss)
tswide_clean %>% features(crab_ho_clinical, unitroot_ndiffs)
tswide_clean %>% features(crab_ho_clinical, unitroot_nsdiffs,.period=12)
tswide_clean %>% gg_tsdisplay(difference(crab_ho_clinical), "partial") 


#### Examine need for a Box-Cox data transformation to reduce heteroskedasticity (i.e variability in the variance of the time-series)
lambda<-guerrero(tswide_clean$crab_ho_clinical, .period=12L)

#### Fit best manual and auto ARIMA on transformed data
arima_crab_ho <- tswide_clean %>% model(
  arima_crab_ho_111 = ARIMA(crab_ho_clinical ~ pdq(1,1,1)),
  arima_crab_ho_011 = ARIMA(crab_ho_clinical ~ pdq(0,1,1)),
  arima_crab_ho_110 = ARIMA(crab_ho_clinical ~ pdq(1,1,0)),
  arima_crab_ho_101 = ARIMA(crab_ho_clinical ~ pdq(1,0,1)),
  stepwise_crab_ho = ARIMA(crab_ho_clinical))

glance(arima_crab_ho) %>% arrange(AICc) %>% select(.model:BIC)

# check the residuals of best models to ensure white noise.
arima_crab_ho  %>% select(arima_crab_ho_011) %>% gg_tsresiduals() 
arima_crab_ho %>% select(arima_crab_ho_101) %>% gg_tsresiduals()

#### Check portmanteau tests of best fit option
augment(arima_crab_ho) %>% 
  filter(.model=='arima_crab_ho_011') %>%
  features(.innov, ljung_box, lag = 10, dof = 3)  # portmanteau test on best manual


```

### 3.1.4 HCA-CRPa clinical infections (*crpa hca clinical*)
```{r message=FALSE, warning=FALSE, include=FALSE}

#### Plot data, ACF, PACF
tswide_clean %>% gg_tsdisplay(crpa_ho_clinical, "partial")

#### Apply a unit root test to check if time-series stationary,  determine number of non-seasonal (ndiffs) and seasonal (nsdiffs) differences required to make stationary, and plot differenced time-series if required.
tswide_clean %>% features(crpa_ho_clinical, unitroot_kpss)
tswide_clean %>% features(crpa_ho_clinical, unitroot_ndiffs) 
tswide_clean %>% features(crpa_ho_clinical, unitroot_nsdiffs,.period=12) 

tswide_clean %>% gg_tsdisplay(difference(crpa_ho_clinical), "partial") 
crpa_ho_clin_diff1 <-  tsibble::difference(tswide_clean$crpa_ho_clinical, lag =1, differences = 1)
tswide_clean %>% gg_tsdisplay(crpa_ho_clin_diff1, "partial")

#### Examine need for a Box-Cox data transformation to reduce heteroskedasticity (i.e variability in the variance of the time-series)
tswide_clean <- tswide_clean %>% dplyr::mutate(crpa_ho_clin_diff1 = replace_na(crpa_ho_clin_diff1, 0))
tswide_clean <- tswide_clean %>% dplyr::mutate(new_crpa_ho_clin_diff1 = crpa_ho_clin_diff1+10)
boxcoxnc(tswide_clean$new_crpa_ho_clin_diff1, method = "sw", lambda2 = 10)


#### Repeat unit root tests and differencing number evaluations on the transformed data
tswide_clean %>% features(crpa_ho_clin_diff1, unitroot_kpss)
tswide_clean %>% features(crpa_ho_clin_diff1, unitroot_ndiffs)
tswide_clean %>% features(crpa_ho_clin_diff1, unitroot_nsdiffs, .period=12)

#### Fit best manual and auto ARIMA
arima_crpa_ho <- tswide_clean %>% model(
  arima_crpa_ho_111 = ARIMA(crpa_ho_clinical ~ pdq(1,1,1)),
  arima_crpa_ho_011 = ARIMA(crpa_ho_clinical ~ pdq(0,1,1)),
  arima_crpa_ho_101 = ARIMA(crpa_ho_clinical ~ pdq(1,0,1)),
  stepwise_crpa_ho = ARIMA(crpa_ho_clinical)) 

glance(arima_crpa_ho) %>% arrange(AICc) %>% select(.model:BIC)

#### Examine residuals of ARIMA to check correct identification
arima_crpa_ho  %>% select(arima_crpa_ho_011) %>% gg_tsresiduals()
arima_crpa_ho %>% select(arima_crpa_ho_101) %>% gg_tsresiduals()
arima_crpa_ho %>% select(stepwise_crpa_ho) %>% gg_tsresiduals() 

#### Check portmanteau tests of best fit option
augment(arima_crpa_ho) %>% 
  filter(.model=='arima_crpa_ho_101') %>%
  features(.innov, ljung_box, lag = 10, dof = 3)  # portmanteau test on best manual
```

## 3.2 Antibiotic and IPC time series
```{r}

# Nb: we use the pre-intervention dataset "ts_p2_pre_icarba" for fitting the baseline ARIMA with exogenous variables (ARIMAX) for carbapenems. For other antibiotics we should use the  "ts_p3_pre_iabs_iunivscr" data
pre_icarba <- ts_p2_pre_icarba %>% as_tsibble(index="date")
pre_iabs <- ts_p3_pre_iabs_iunivscr %>% as_tsibble(index="date")



```
## 3.3 Influential infection time-series
These time-series are expected to be related to antibiotic use in the population. Bidirectional relationships may exist but we hypothesised that hospital-wide use of important antibiotic exposures driving carbapenem resistance may be responsive to the following:

* All gram-negative bloodstream isolates (BSIs) (*gmneg_all_bsi*)
* Carbapenem and cephalosporin sensitive E.coli BSI (*ecoliccs_all_bsi*)
* 3rd generation cephalosporins resistant E.coli BSI (*ecoli3gcr_all_bsi*)

### 3.3.1 Gram-Neg BSIs
```{r}
### Carbapenems######################

# Prepare differenced time-series
pre_icarba$gmneg_diff <-  tsibble::difference(pre_icarba$gmneg_all_bsi, lag =1, differences = 1)
pre_icarba$carbapenem_diff <-  tsibble::difference(pre_icarba$carbapenems, lag =1, differences = 1)
pre_icarba_ed2 <- pre_icarba [2:72,] %>% as_tsibble(index="date")
pre_icarba_ed4 <- pre_icarba [4:72,] %>% as_tsibble(index="date")

# Explore cross-correlations with raw data
CCF(pre_icarba, carbapenems, gmneg_all_bsi, lag_max=24) %>% autoplot(level=95)

# Explore cross-correlations with differenced data
CCF(pre_icarba_ed2, carbapenem_diff, gmneg_diff, lag_max=24)%>% autoplot(level=95)

# Identify an ARIMAX model for carbapenems 
carba_pre1 <- pre_icarba %>% model(ARIMA(carbapenems  ~ -1 + pdq(0,1,1)+PDQ(0,0,0)+lag(gmneg_all_bsi,2)))
report(carba_pre1)
carba_pre1 %>%  gg_tsresiduals() 

carba_pre2 <- pre_icarba_ed2 %>% model(ARIMA(carbapenems  ~ -1 + pdq(0,1,1)+PDQ(0,0,0)+lag(gmneg_diff,2)))
report(carba_pre2)
carba_pre2 %>%  gg_tsresiduals() 

# Looking at the correlation between gm neg BSI and differenced gm neg BSI (y-axis)
gmneg_bsi<- pre_icarba$gmneg_all_bsi[1:71]
differenced_gm_negbsi <- pre_icarba$gmneg_diff [2:72]
gmneg <- cbind(gmneg_bsi,differenced_gm_negbsi )
gmneg <-gmneg %>% as_tibble()
cor(gmneg$gmneg_bsi, gmneg$differenced_gm_negbsi)
ggplot(gmneg, aes(gmneg_bsi,differenced_gm_negbsi))+geom_point()

```
# 4. Interactions and VARs

# 5. GAMs

# 6. MARS



```{r}
plot(cars)
```

